---
title: "The title"
author:
- name: First Author
  affiliation: '1'
  corresponding: true
  address: Postal address
  email: my@email.com
  role:
  - Conceptualization
  - "Writing - Original Draft Preparation"
  - "Writing - Review & Editing"
- name: "Ernst-August Doelle"
  affiliation: '1,2'
  role:
  - "Writing - Review & Editing"
  - Supervision
shorttitle: Title
affiliation:
- id: '1'
  institution: "Wilhelm-Wundt-University"
- id: '2'
  institution: Konstanz Business School
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: "One or two sentences providing a **basic introduction** to the field,  comprehensible
  to a scientist in any discipline.\nTwo to three sentences of **more detailed background**,
  comprehensible  to scientists in related disciplines.\nOne sentence clearly stating
  the **general problem** being addressed by  this particular study.\nOne sentence
  summarizing the main result (with the words \"**here we show**\" or their equivalent).\nTwo
  or three sentences explaining what the **main result** reveals in direct comparison
  to what was thought to be the case previously, or how the  main result adds to previous
  knowledge.\nOne or two sentences to put the results into a more **general context**.\nTwo
  or three sentences to provide a **broader perspective**, readily comprehensible
  to a scientist in any discipline.\n\n<!-- https://tinyurl.com/ybremelq -->\n"
keywords: keywords
wordcount: X
bibliography: "r-references.bib"
floatsintext: false
linenumbers: true
draft: false
mask: false
figurelist: false
tablelist: false
footnotelist: false
classoption: man
---


```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(psych)
library(modelr)
library(glue)
library(here)
library(quanteda)
library(tmcn)
library(janitor)
library(GGally)
library(lme4)
library(stringr)
library(lmerTest)  # For p-values in lme4 models
library(skimr)     # For quick and nice summaries
library(performance) # For model diagnostics
library(ggvenn)
library(purrr)
source(here("helper.R"))
library("papaja")
r_refs("r-references.bib")

```

# Introduction

# Study 1: 

## Methods
In Study 1, we aimed to replicate (Samuelson 1999) by collecting judgment ratings on the first 300 words learned from the MCDI US English across the same three dimensions of solidity, count, and mass syntax, and shape-based categorization. The study, however, collected these ratings on the first 300 words learned by children in sixteen languages including English. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

english_demo1 <- read_csv(here("data/demographics/first-sample-english", "1.csv"))
english_demo2 <- read_csv(here("data/demographics/first-sample-english", "2.csv"))
english_demo3 <- read_csv(here("data/demographics/first-sample-english", "3.csv")) %>%
  filter(!is.na(Age)) %>% mutate(Age = as.character(Age))
english_demo4 <- read_csv(here("data/demographics/first-sample-english", "4.csv")) %>%
  filter(!is.na(Age)) %>% mutate(Age = as.character(Age))
english_demo5 <- read_csv(here("data/demographics/first-sample-english", "5.csv"))

english_sample1_demo <- bind_rows(english_demo1, english_demo2, english_demo3, english_demo4, english_demo5) %>%
  clean_names() %>%
  mutate(language = "English", sample = "1") %>%
  mutate(age = as.numeric(age)) %>% 
  drop_na(age)

# %>% filter(age >= 18 & age <= 60)

english_sample1_demo %>% skim()
english_sample1_demo %>% select(age) %>% filter(!is.na(age)) %>%
summarize(meanage = mean(age), sd = sd(age), max = max(age), min = min(age))

```

## Participants
377 English-speaking adults recruited online from the Prolific platform (mean age = 41, sd = 14) were asked to rate words in three dimensions: solidity (solid, non-solid, unclear), count and mass noun syntax (count, mass, unclear), and organizing feature (shape, color, material, none of these). 

## Material
Three hundred nouns were sourced from MCDI data on the WordBank repository of 16 languages that belonged to seven linguistic families: (LIST OF LANGUAGES). The nouns were chosen based on the earliest age of acquisition in the dataset, up to a maximum of 300 nouns (CITE). A Total of 23355 nouns overall, with 664 after removing duplicates. (Figure to show how much overlap between languages). With the Semantic categories breakdown.
Because we only recruited English-speaking participants, Nouns were presented in English and mapped across languages via conceptual mappings (“unilemmas”, Frank et al., 2021; N = ~664). A “unilemma” is the concept behind the noun expressed in the English language (“dog” and “كلب” both map to the same underlying concept of “dog”, which is used as the unilemma in this case). 

## Procedure
The experiment started with familiarization trials in which participants were shown examples of count and mass nouns, solid objects, and non-solid objects. They were also familiarized with what a category-organizing or characterizing feature means, as in Table [2]. Then the test trials followed the same wording and structure. Every dimension (solidity, count-mass syntax, and category organizing feature) constituted a block, and every participant rated 20 words per block, so there were 60 test trials in total. Familiarization trials, test blocks, and words within a block were all randomized across participants. No information about the semantic category of words was given to participants, nor were the words grouped by their semantic category. Participants were only allowed to choose one feature in all questions. 

## Data analysis

## Results
```{r echo=FALSE, message=FALSE, warning=FALSE}
english_sample1 <- read_csv(here("data/predictors","english_sample1_props.csv")) 
english_sample1 <- english_sample1 %>% filter(!is.na(response)) %>% distinct()
english_sample1_raw <- read_csv(here("data/predictors","english_sample1_raw.csv")) 
# Apply the mapping function in helper.R to the response column
english_sample1$response_numeric <- mapply(map_categories, english_sample1$response, english_sample1$block)
# more 2 participants -- 21 words less than 8 ratings
```

```{r }
english_sample1 |>
  filter(block == "category_organization") |>
ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language) +
  labs(
    title = "Density of Proportions by Language",
    x = "Proportion Value",
    y = "Density (words)"
  ) +
  theme_minimal()

english_sample1 |>
  filter(block == "solidity") |>
ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language) +
  labs(
    title = "Density of Proportions by Language",
    x = "Proportion Value",
    y = "Density (words)"
  ) +
  theme_minimal()

english_sample1 |>
  filter(block == "count_mass") |>
ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language) +
  labs(
    title = "Density of Proportions by Language",
    x = "Proportion Value",
    y = "Density (words)"
  ) +
  theme_minimal()
```

```{r correspondence_eng_1}
correspondence <- english_sample1 |> filter (matching_response %in% c("solid", "count" ,"shape"))

correspondence |>
  ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language) +
  labs(
    title = "Density of Proportions by Language",
    x = "Proportion Value",
    y = "Density (words)"
  ) +
  theme_minimal()

```

```{r venn_eng_1}
#Venn Diagram

threshold <- 0.7

category_english_sample1 <- english_sample1 |>
  filter(response %in% c("solid","count noun","shape")) |>
  mutate(over_threshold = proportion > threshold) |>
  select(-block, -totcount, -count, -proportion) |> 
  pivot_wider(names_from = "response", values_from = "over_threshold", 
              values_fill = FALSE) |>
  rename(count_noun = `count noun`)


venns <- category_english_sample1 |>
  group_by(language) |>
  nest() |>
  mutate(plot = map(data, ~ggvenn(.x, c("count_noun","solid","shape"),
                                  show_percentage = FALSE, 
                                  text_size = 1,
                                  set_name_size = 2)
                    + ggtitle(language))) |>
  pull(plot)

grid.arrange(grobs = venns, ncol = 4)

```

We collected a total of 20,130 ratings across 377 participants for 664 words in three blocks (solidity, count-mass syntax, and category-organizing feature). After removing responses with no answer, we retained 19,635 ratings. Each word received an average of between 8 and 10 ratings per block. Included in the supplementary materials are the demographic details of the participants (Table 1) and the distribution of ratings across words and blocks (Figure 1).

Across languages, nouns were predominantly rated as solid (M = 0.72, SD = 0.32) and count (M = 0.68, SD = 0.29), but were less reliably judged as being organized by shape (M = 0.31, SD = 0.24) as can be seen in figure \ref{fig:correspondence_eng_1}. The venn diagram in figure \ref{fig:venn_eng_1} shows categories of words as solid, count noun, and shape_characeterized by using a cutoff of 80% i.e. a word that is rated by at least 80% of the participants to be a solid is categorized as solid. The venn diagram shows that a large number of nouns were consistently rated as solid and count nouns, with more overlap between these two dimensions those with the minority of words rated as referring to objects organized by shape. 

Ratings were very consistent across languages, suggesting limited variation in lexical statistics. To test for distributional differences across languages, we conducted a permutation test (see Methods). The results showed that the observed density curves for each language fell within the null distribution generated by permuting language labels, indicating no significant differences in lexical statistics across languages (Figure 2).

```{r}
#permutation test

# Preparation: isolate one response (e.g. “shape”)
english_sample1_shape_df <- english_sample1 %>%
  filter(response == "shape") %>%
  replace_na(list(proportion = 0)) %>%
  select(uni_lemma, language, proportion)

english_sample1_shape_df %>% glimpse()


```

2. Compute the observed density curve for each language

```{r}
# 2.A. Decide on a grid of proportion‐values between 0 and 1
grid_x <- seq(0, 1, length.out = 100)

# 2.B. A helper function that, given a vector of proportions, returns its density 
#      evaluated at the grid points.
get_density_on_grid <- function(x_values, grid_x) {
  # Remove any NAs (we replaced above)
  x_vals <- x_values[!is.na(x_values)]
  
  # If a language has fewer than 2 unique points, density() might fail—so 
  # we require at least 2 non‐identical values to form a kernel density.
  if(length(unique(x_vals)) < 2) {
    # Return a vector of zeros or NA so that it’s the “degenerate” distribution.
    return(rep(0, length(grid_x)))
  }
  
  d <- density(x_vals, from = 0, to = 1, n = length(grid_x))
  # d$y is the estimated density at each of the grid points; d$x is almost equal to grid_x
  # To be safe, we’ll interpolate:
  approx(d$x, d$y, xout = grid_x, rule = 2)$y
}

# 2.C. Compute the *observed* densities for each language
obs_density_df <- english_sample1_shape_df %>%
  group_by(language) %>%
  summarize(
    dens = list(get_density_on_grid(proportion, grid_x)),
    .groups = "drop"
  ) %>%
  # Turn the list‐column “dens” into one row per (language, grid_x)
  unnest_longer(dens, values_to = "density") %>%
  group_by(language) %>%
  mutate(x = grid_x) %>%
  ungroup()

```

3. Build a Null Distribution by Permuting “language” labels

```{r}
set.seed(123)  # for reproducibility
n_perm <- 500

# A helper: given a data frame, shuffle language labels and recompute densities
one_permuted_densities <- function(df, grid_x) {
  shuffled <- df %>%
    mutate(language = sample(language))  # randomly shuffle language labels
  
  # now recompute densities by “fake” language grouping
  perm_df <- shuffled %>%
    group_by(language) %>%
    summarize(
      dens = list(get_density_on_grid(proportion, grid_x)),
      .groups = "drop"
    ) %>%
    unnest_longer(dens, values_to = "density") %>%
    group_by(language) %>%
    mutate(x = grid_x) %>%
    ungroup()
  
  # We want a tidy format: (language, x, density_perm_i)
  perm_df
}

# 3.A. Run all permutations, storing results in a single data frame
perm_densities <- map_dfr(seq_len(n_perm), function(i) {
  one_permuted_densities(english_sample1_shape_df, grid_x) %>%
    mutate(perm = i)
})

# perm_densities now has columns: language, x, density, perm
# For each (language, x), we have n_perm density values.

```


```{r}
null_envelopes <- perm_densities %>%
  group_by(language, x) %>%
  summarize(
    lower = quantile(density, 0.025),
    upper = quantile(density, 0.975),
    .groups = "drop"
  )

```


5. Plot Observed vs. Null for Each Language
```{r}
ggplot() +
  # NULL envelope bands (gray ribbons)
  geom_ribbon(
    data = null_envelopes,
    aes(x = x, ymin = lower, ymax = upper),
    fill = "gray80", alpha = 0.5
  ) +
  # OBSERVED density curves (bold lines)
  geom_line(
    data = obs_density_df,
    aes(x = x, y = density),
    color = "blue", size = 1
  ) +
  facet_wrap(~ language, scales = "free_y") +
  labs(
    title = "Observed 'shape' density vs. Null Envelope by Language",
    x = "Proportion of 'shape' choices",
    y = "Density"
  ) +
  theme_minimal()

```

```{r}
# Quantify Difference: KS‐Stat on Observed vs. Null

# First compute the median permuted density at each (language, x)
median_perm <- perm_densities %>%
  group_by(language, x) %>%
  summarize(median_density = median(density), .groups = "drop")

# Join with observed
diff_df <- obs_density_df %>%
  left_join(median_perm, by = c("language", "x")) %>%
  mutate(diff_from_median = abs(density - median_density))

# Integrate difference over x (approximate area between curves)
area_diff <- diff_df %>%
  group_by(language) %>%
  summarize(
    area = sum(diff_from_median) * (1 / (length(grid_x) - 1)),  # Riemann approx
    .groups = "drop"
  )

# area now is a single “distance” metric for each language.
print(area_diff)

```

```{r}
# how much correspondence there is between the three dimensions 
correspodence_data <-  english_sample1 %>% select(-c(language)) %>%
  distinct() %>%
  filter(response %in% c("solid", "shape", "count noun")) %>%  # Filter only the desired responses
  select(word, response, proportion) %>%  # Keep only relevant columns
  pivot_wider(names_from = response, values_from = proportion, values_fill = 0)  # Reshape

correspodence_data %>% select(`count noun`, shape, solid) %>% corr.test()

```

```{r}
#inter-rater agreement in the word level: by word-level consensus

get_mode <- function(v) {
  uniq_v <- unique(na.omit(v))
  uniq_v[which.max(tabulate(match(v, uniq_v)))]
}


consensus_data <- english_sample1_raw %>%
  group_by(block, uni_lemma, word) %>%
  summarise(
    mode_rating = get_mode(response_numeric),
    agreement_score = (mean(response_numeric == mode_rating, na.rm = TRUE) -0.009) * 100,
    sd_rating = sd(response_numeric, na.rm = TRUE)
  ) 

ggplot(consensus_data, aes(x = sd_rating, y = agreement_score, color = block)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Agreement Score vs. Rating Standard Deviation", x = "Standard Deviation of Ratings", y = "Agreement Score (%)") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3")

```

# test-retest
reliability

```{r }
english_sample2 <- read_csv(here("data/predictors","english_sample2_props.csv"))
english_sample2 <- english_sample2 %>% filter(!is.na(response)) %>% distinct()
# more 15 participants -- 100 words less than 8 ratings

korean_sample1 <- read_csv(here("data/predictors","korean_sample1_props.csv")) 
korean_sample1 <- korean_sample1 %>% filter(!is.na(response)) %>% distinct()
# more 60 participants -- 241 words -- pending

korean_sample2 <- read_csv(here("data/predictors","korean_sample2_props.csv"))
korean_sample2 <- korean_sample2 %>% filter(!is.na(response)) %>% distinct()
# no more participants needed


# english_sample1 %>% filter(totcount < 8, block == "category_organization") %>% select(word) %>% distinct() %>% jsonlite::toJSON()

korean_sample1 %>% filter(totcount > 8) %>% select(theword) %>% distinct() %>% jsonlite::toJSON()

```

```{r }
english_sample2 |>
  filter(block == "category_organization") |>
ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language) +
  labs(
    title = "Density of Proportions by Language",
    x = "Proportion Value",
    y = "Density (words)"
  ) +
  theme_minimal()
```

```{r}
eng_samp1_shape <- english_sample1 %>% filter(response == "shape") %>% select(word, proportion)
eng_samp2_shape <- english_sample2 %>% filter(response == "shape") %>% select(word, proportion)

# Merge the data frames based on 'word_id'
merged_data <- merge(eng_samp1_shape, eng_samp2_shape, by = "word", suffixes = c(".sample1", ".sample2")) %>% distinct()

# Calculate the correlation between proportion values
correlation_score <- cor(merged_data$proportion.sample1, merged_data$proportion.sample2, use = "complete.obs") # Handles missing values

# Print the correlation score
print(correlation_score)
```

# Study 2:
## Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
