
# Study 1

```{r message=FALSE, warning=FALSE}
english_sample1 <- read_csv(here("data/predictors","english_sample1_props.csv")) 
english_sample1 <- english_sample1 %>% filter(!is.na(response)) %>% distinct()
english_sample1_raw <- read_csv(here("data/predictors","english_sample1_raw.csv")) 
# Apply the mapping function in helper.R to the response column
english_sample1$response_numeric <- mapply(map_categories, english_sample1$response, english_sample1$block)
# more 2 participants -- 21 words less than 8 ratings

correspondence <- english_sample1 |> filter (matching_response %in% c("solid", "count" ,"shape"))

```

```{r , echo=TRUE}
# how much correspondence there is between the three dimensions 
correspodence_data <-  english_sample1 %>% select(-c(language)) %>%
  distinct() %>%
  filter(response %in% c("solid", "shape", "count noun")) %>%  # Filter only the desired responses
  select(word, response, proportion) %>%  # Keep only relevant columns
  pivot_wider(names_from = response, values_from = proportion, values_fill = 0)  # Reshape

correspodence_data %>% select(`count noun`, shape, solid) %>% corr.test()

```

```{r agreement_plot, include=TRUE}
consensus_data <- english_sample1_raw %>%
  group_by(block, uni_lemma, word) %>%
  summarise(
    mode_rating = get_mode(response_numeric),
    agreement_score = (mean(response_numeric == mode_rating, na.rm = TRUE) -0.009) * 100,
    sd_rating = sd(response_numeric, na.rm = TRUE)
  )

ggplot(consensus_data, aes(x = sd_rating, y = agreement_score, color = block)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Agreement Score vs. Rating Standard Deviation", x = "Standard Deviation of Ratings", y = "Agreement Score (%)") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3")
```

# Study 2
```{r }
english_sample2 <- read_csv(here("data/predictors","english_sample2_props.csv"))
english_sample2 <- english_sample2 %>% filter(!is.na(response)) %>% distinct()
english_sample2_raw <- read_csv(here("data/predictors","english_sample2_raw.csv"))
# Apply the mapping function in helper.R to the response column
english_sample2$response_numeric <- mapply(map_categories, english_sample2$response, english_sample2$block)

# more 1 or 2 participants -- 29 words less than 8 ratings

# english_sample1 %>% filter(totcount < 8, block == "category_organization") %>% select(word) %>% distinct() %>% jsonlite::toJSON()

#english_sample2 %>% filter(totcount < 8, block == "category_organization") %>% select(word) %>% distinct() %>% jsonlite::toJSON()

```

# Reliability analysis 

```{r }
eng_samp1_shape <- english_sample1 %>% filter(response == "shape") %>% select(word, proportion)
eng_samp2_shape <- english_sample2 %>% filter(response == "shape") %>% select(word, proportion)

# Merge the data frames based on 'word_id'
merged_data <- merge(eng_samp1_shape, eng_samp2_shape, by = "word", suffixes = c(".sample1", ".sample2")) %>% distinct()
```

## split-half estimate of reliability
through simple pearson correlation 
```{r}
# Calculate the correlation between proportion values
correlation_score <- cor(merged_data$proportion.sample1, merged_data$proportion.sample2, use = "complete.obs") # Handles missing values

# Print the correlation score
print(paste("correlation between shape proportions in 1A VS 1B is: ", correlation_score))

```

```{r }
ggplot(merged_data, aes(x = proportion.sample1, y = proportion.sample2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Item-Level Agreement: Shape Ratings in Study 1A vs 1B")
```

## overall distribution
```{r }
# Load your data (adjust file paths as needed)
study1a_data <- read_csv(here("data/predictors","study1a_data.csv"))

#fix names of columns
study1a_data <- study1a_data %>%
  rename(participant_id = `_id`)

study1b_data <-  read_csv(here("data/predictors","study1b_data.csv"))
study1b_data <- study1b_data %>%
  rename(participant_id = `_id`)


# 1. DATA PREPARATION
# ===================
study1a_data<- study1a_data %>% 
    select(uni_lemma, language, shape_rating) %>%
    mutate(study = "1A")

study1b_data <- study1b_data %>% 
    select(uni_lemma, language, shape_rating) %>%
    mutate(study = "1B")

# Combine datasets with study identifier
combined_data <- bind_rows( 
  study1b_data, study1a_data
   ) %>%
  filter(!is.na(shape_rating)) %>%  # Remove missing values
mutate(shape_rating = as.numeric(shape_rating)) %>%
  distinct()  # Remove duplicates

combined_data %>% filter(study == "1A")
combined_data %>% filter(study == "1B")

# Create wide format for reliability analysis
ratings_wide <- combined_data %>%
  pivot_wider(names_from = study, 
              values_from = shape_rating,
              names_prefix = "Study_") %>%
  filter(!is.na(Study_1A) & !is.na(Study_1B))  # Only words rated in both studies

```

Use a Kolmogorov–Smirnov (KS) test to see if the two distributions differ significantly.
```{r }
ks.test(combined_data$shape_rating[combined_data$study == "1A"],
        combined_data$shape_rating[combined_data$study == "1B"])

```

A classic example of low item-level agreement but high aggregate-level agreement.

It doesn’t necessarily mean the measure is unreliable — instead, it may suggest:

That there's measurement noise at the item level (e.g., variability in how different raters interpret the words).

Or that your construct is more meaningful at the group level — i.e., shape ratings are better thought of as a property of categories or populations, not precise values for individual words.

In psychometrics, this distinction often arises between individual-level reliability (e.g., inter-rater reliability for specific items) and scale-level consistency (e.g., internal consistency across a battery of items).

## ICC

cor()) tells how well rank order is preserved. Intraclass Correlation Coefficient (ICC) tells how much of the variance in scores is due to between-item differences vs. noise. Useful for assessing rater consistency or measurement reliability.
```{r }
library(irr)
icc_result <- icc(data.frame(merged_data$proportion.sample1, merged_data$proportion.sample2), model = "twoway", type = "consistency", unit = "single")
print(icc_result)

```

## Bland-Altman Plot
visualize agreement between two measurements (especially useful when comparing two raters or samples). shows the mean difference between ratings and any bias or systematic shift across rating ranges

```{r }
if (!requireNamespace("BlandAltmanLeh", quietly = TRUE)) {
      install.packages("BlandAltmanLeh")
}
library(BlandAltmanLeh)

bland.altman.plot(merged_data$proportion.sample1, merged_data$proportion.sample2)

```

## Prophecy: spearman-brown prediction formula 
``` {r}
predicted_reliability <- (2 * correlation_score) / (1 + correlation_score)
```
we examined the measurement properties by measuring reliability, then examine the 


## Using pearson correlation between two halves of the data (split-half reliability)

## Internal consistency (Cronbach's alpha) for each dimension in each study

## Item-Level Error Inspection

```{r}
merged_data$abs_diff <- abs(merged_data$proportion.sample1 - merged_data$proportion.sample2)
summary(merged_data$abs_diff)
# Identify items with the largest discrepancies
top_discrepancies <- merged_data %>% arrange(desc(abs_diff)) %>%
  head(15)
print(top_discrepancies)
```


