---
title: "Reverse causality and omitted variables in the shape-AoA association (Study 4)"
author: "Lexical statistics project"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(lme4)
library(lmerTest)
library(readxl)
library(janitor)
```

This document addresses the reverse-causality and omitted-variable concern for
the Study 4 result (the shape rating predicts earlier AoA): adult raters may
judge a noun shape-organized partly *because* it is a familiar, early, highly
imageable word, rather than the lexical statistic driving acquisition. The
check is to refit the Table 3 model with covariates that index the
"familiar early word" pathway: adult imageability and familiarity norms
(Glasgow Norms; Scott et al., 2019), adult English AoA norms (Kuperman,
Stadthagen-Gonzalez, & Brysbaert, 2012), and the English-anchored item norms
used in the Wordbank prediction literature (babiness, valence, arousal;
Braginsky et al., 2019). If the shape coefficient survives these covariates,
a pure reverse-causality account becomes harder to maintain; if it does not,
the association cannot be separated from adult lexical familiarity. Because
Braginsky et al.'s language-specific covariates (MLU, phoneme counts) exist
for only four of our fourteen languages, we use their English-anchored
item-level norms, which enter exactly the way concreteness already does in
the manuscript.

# Data

```{r data-load}
# Primary Study 4 analysis set (one row per unilemma-by-language observation).
ratings_aoa <- read_csv(here("data/predictors", "ratings_aoa.csv"),
                        show_col_types = FALSE)

# Unilemma cleaning for matching against single-word English norms:
# lowercase, strip disambiguating parentheticals ("chicken (animal)"),
# and take the first alternative of slash-separated forms ("soda/pop").
clean_lemma <- function(x) {
  x %>% str_to_lower() %>%
    str_remove("\\s*\\(.*\\)\\s*$") %>%
    str_split_fixed("/", 2) %>% .[, 1] %>%
    str_trim()
}
ratings_aoa <- ratings_aoa %>% mutate(match_word = clean_lemma(uni_lemma))

# Kuperman et al. (2012) adult AoA ratings (downloaded from osf.io/d7x6q).
kuperman <- read_xlsx(here("data/predictors/external",
                           "kuperman_aoa_with_pos.xlsx")) %>%
  clean_names() %>%
  transmute(match_word = str_to_lower(word),
            adult_aoa  = as.numeric(rating_mean)) %>%
  filter(!is.na(adult_aoa)) %>%
  distinct(match_word, .keep_all = TRUE)

# Glasgow Norms (Scott et al., 2019): imageability and familiarity
# (downloaded from the JASP data library mirror of the BRM supplement).
glasgow <- read_csv(here("data/predictors/external", "glasgow_norms.csv"),
                    show_col_types = FALSE) %>%
  clean_names() %>%
  transmute(match_word   = clean_lemma(word),
            imageability = as.numeric(m_imag),
            familiarity  = as.numeric(m_fam)) %>%
  group_by(match_word) %>%
  summarise(across(c(imageability, familiarity), mean), .groups = "drop")

# Braginsky et al. (2019) English-anchored item norms, extracted from the
# aoa-prediction pipeline (uni_model_data.RData), English (American) rows.
braginsky <- read_csv(here("data/predictors/external",
                           "braginsky_uni_predictors.csv"),
                      show_col_types = FALSE) %>%
  filter(language == "English (American)") %>%
  transmute(match_word = clean_lemma(uni_lemma),
            babiness, valence, arousal) %>%
  distinct(match_word, .keep_all = TRUE)

model_df <- ratings_aoa %>%
  left_join(kuperman,  by = "match_word") %>%
  left_join(glasgow,   by = "match_word") %>%
  left_join(braginsky, by = "match_word") %>%
  distinct(language, uni_lemma, .keep_all = TRUE)

coverage <- tibble(
  covariate = c("adult AoA (Kuperman)", "imageability (Glasgow)",
                "familiarity (Glasgow)", "babiness (Braginsky)"),
  n_unilemmas = c(
    n_distinct(model_df$uni_lemma[!is.na(model_df$adult_aoa)]),
    n_distinct(model_df$uni_lemma[!is.na(model_df$imageability)]),
    n_distinct(model_df$uni_lemma[!is.na(model_df$familiarity)]),
    n_distinct(model_df$uni_lemma[!is.na(model_df$babiness)])),
  n_total = n_distinct(model_df$uni_lemma))
knitr::kable(coverage, caption = "Norm coverage of the Study 4 unilemma set.")
```

# Models

Each augmented model is compared against the Table 3 baseline refit on the
same rows (complete cases for that covariate set), so coefficient changes
reflect the added covariates rather than the changed sample.

```{r models}
z <- function(x) as.numeric(scale(x))

fit_pair <- function(df, extra_terms, label) {
  covs <- c("logfreq", "concreteness", "solid", "mean_shape", "count_noun",
            extra_terms)
  d <- df %>%
    select(uni_lemma, aoa, category, language, all_of(covs)) %>%
    drop_na() %>%
    mutate(across(all_of(covs), z)) %>%
    distinct()
  base_form <- aoa ~ logfreq + concreteness + solid + mean_shape +
    count_noun + language + (1 | category) + (1 | uni_lemma)
  aug_form  <- update(base_form,
                      paste(". ~ . +", paste(extra_terms, collapse = " + ")))
  fit_b <- lmerTest::lmer(base_form, data = d)
  fit_a <- lmerTest::lmer(aug_form,  data = d)
  cb <- coef(summary(fit_b))["mean_shape", ]
  ca <- coef(summary(fit_a))["mean_shape", ]
  tibble(
    model      = label,
    n_rows     = nrow(d),
    n_unilemma = n_distinct(d$uni_lemma),
    shape_b_baseline = cb[["Estimate"]], shape_p_baseline = cb[["Pr(>|t|)"]],
    shape_b_augmented = ca[["Estimate"]], shape_p_augmented = ca[["Pr(>|t|)"]],
    extra = list(coef(summary(fit_a))[extra_terms, , drop = FALSE]))
}

res <- bind_rows(
  fit_pair(model_df, c("imageability", "familiarity"),
           "+ imageability + familiarity"),
  fit_pair(model_df, "adult_aoa", "+ adult English AoA"),
  fit_pair(model_df, c("babiness", "valence", "arousal"),
           "+ babiness + valence + arousal"),
  fit_pair(model_df, c("imageability", "familiarity", "adult_aoa",
                       "babiness", "valence", "arousal"),
           "kitchen sink (all six)"))

res %>%
  select(-extra) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  knitr::kable(caption = paste(
    "Shape coefficient (months of AoA per SD of shape rating) before and",
    "after adding the covariates, on identical rows."))
```

```{r extra-coefs}
# Coefficients of the added covariates in each augmented model.
for (i in seq_len(nrow(res))) {
  cat("\n##", res$model[i], "\n")
  print(round(res$extra[[i]], 4))
}
```

# Range restriction of the adult norms in this item set

```{r range-restriction}
rr <- bind_rows(
  tibble(norm = "adult AoA (Kuperman)",
         sd_our_items = sd(model_df %>% distinct(uni_lemma, adult_aoa) %>%
                             pull(adult_aoa), na.rm = TRUE),
         sd_full_norms = sd(kuperman$adult_aoa)),
  tibble(norm = "imageability (Glasgow)",
         sd_our_items = sd(model_df %>% distinct(uni_lemma, imageability) %>%
                             pull(imageability), na.rm = TRUE),
         sd_full_norms = sd(glasgow$imageability, na.rm = TRUE)),
  tibble(norm = "familiarity (Glasgow)",
         sd_our_items = sd(model_df %>% distinct(uni_lemma, familiarity) %>%
                             pull(familiarity), na.rm = TRUE),
         sd_full_norms = sd(glasgow$familiarity, na.rm = TRUE))) %>%
  mutate(sd_ratio = sd_our_items / sd_full_norms,
         across(where(is.numeric), ~ round(.x, 3)))
knitr::kable(rr, caption = paste(
  "Spread of the adult norms among our early nouns relative to the full",
  "norm sets: early nouns are a restricted slice of the adult lexicon."))
```

# Results

*(Drafted as manuscript text; update the numbers from the tables above if
the data change.)*

```{r results-text, echo=FALSE, results="asis"}
fmt <- function(b, p) sprintf("$B$ = %.2f, $p$ = %.3g", b, p)
r1 <- res[1, ]; r2 <- res[2, ]; r3 <- res[3, ]; r4 <- res[4, ]
cat(sprintf(
  paste(
    "A remaining concern is reverse causality: adults may rate familiar,",
    "early, highly imageable words as shape-organized, rather than shape",
    "organization driving acquisition. We therefore refit the primary model",
    "adding adult norms that index that pathway. Adding Glasgow imageability",
    "and familiarity norms (matched for %d unilemmas) left the shape",
    "coefficient essentially unchanged (%s, against %s in the baseline",
    "refit on the same rows). Adding adult English AoA (Kuperman et al.,",
    "2012; %d unilemmas), itself a strong predictor of child AoA, gave %s",
    "for shape. With all six adult covariates entered together (imageability,",
    "familiarity, adult AoA, babiness, valence, arousal), the shape",
    "coefficient was %s. On the smaller subset with babiness coverage",
    "(%d unilemmas), shape was not reliable with or without the added",
    "covariates (%s augmented; %s baseline on the same rows), so that null",
    "reflects the reduced item set rather than the covariates. These fits do",
    "not identify the causal direction, but they are inconsistent with the",
    "specific alternative on which the shape rating merely proxies adult",
    "familiarity or the adult accessibility of early words.\n"),
  r1$n_unilemma,
  fmt(r1$shape_b_augmented, r1$shape_p_augmented),
  fmt(r1$shape_b_baseline,  r1$shape_p_baseline),
  r2$n_unilemma,
  fmt(r2$shape_b_augmented, r2$shape_p_augmented),
  fmt(r4$shape_b_augmented, r4$shape_p_augmented),
  r3$n_unilemma,
  fmt(r3$shape_b_augmented, r3$shape_p_augmented),
  fmt(r3$shape_b_baseline,  r3$shape_p_baseline)))
```

Two caveats stay open and are worth stating in any response letter. First,
the norms are English ratings attached to unilemmas, so they are constant
across languages the way concreteness is; they control the item-level
pathway, not language-specific familiarity. Second, our nouns are a
restricted slice of the adult lexicon (see the range-restriction table), so
the adult covariates enter with less variance than they carry in the full
norm sets, and their coefficients here should not be compared with estimates
from unrestricted vocabularies.
