---
title: "Supplementary materials: Lexical statistics and the shape bias"
author: "Authors masked for review"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: pdflatex
    toc: false
header-includes:
  - \usepackage{graphicx}
  - \usepackage{booktabs}
  - \usepackage{threeparttable}
  - \usepackage{float}
  - \usepackage{etoolbox}
  - \AtBeginEnvironment{table}{\small\setlength{\tabcolsep}{4.5pt}}
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.path   = "data/figures/supplement-",
  fig.align  = "center",
  dev        = "pdf",
  dpi        = 300,
  out.width  = "0.98\\textwidth",
  fig.width  = 5.4,
  fig.height = 3.0
)

library(here)
library(tidyverse)
library(papaja)
library(psych)
library(eulerr)
library(gridExtra)
library(janitor)
library(lme4)
library(lmerTest)

source(here("helper.R"))

euler_fills <- c("grey83", "steelblue4", "firebrick3")
euler_panel <- function(df, title) {
  fit <- euler(as.data.frame(df))
  plot(fit,
       fills = list(fill = euler_fills, alpha = 0.55),
       edges = list(col = "grey30", lwd = 0.5),
       labels = list(fontsize = 4),
       quantities = list(fontsize = 3.5),
       adjust_labels = TRUE,
       main = list(label = title, fontsize = 5))
}
```

This document supplements the main manuscript. Contents: (S1) trial wording for the online rating tasks; (S2) Study 1A lexicon composition, item-count reconciliation across studies, demographics, rating distributions, and missing-identifier sensitivity; (S3) full 16-language density and Euler grids; (S4) cross-sample shape correlations; (S5) Korean age-of-acquisition models; (S6) Study 2A strict- vs. broad-shape coding; (S7) pooled English rater counts by procedure; (S8) leave-one-language-out refits of the Study 4 model; (S9) Bayesian posterior and power sensitivity for the shape--AoA coefficient.

# S1. Trial wording (Studies 1A and 2B)

Interactive browser implementations of the rating tasks accompany the Open Materials for this manuscript. The procedures match the main text.

**Study 1A.** Participants completed three blocks in random order (solidity, count–mass syntax, category-organizing feature), 20 nouns per block. Solidity and count–mass trials presented a single unilemma as the prompt. The organizing-feature trial asked whether the word *belongs to a category that is organized by* shape, color, material, or none of these (single forced choice). Words were not grouped by semantic category.

**Study 2B.** Trial structure matched Study 1A for the corresponding blocks. Instructions used the clarified organizing-feature wording described in the main text.

# S2. Study 1A: Lexicon composition, demographics, and rating diagnostics

```{r load-eng1}
english_sample1 <- read_csv(here("data/predictors", "english_sample1_props.csv"),
                           show_col_types = FALSE) %>%
  filter(!is.na(response)) %>%
  distinct()
threshold <- 0.8

langs_16 <- english_sample1 %>%
  distinct(language) %>%
  arrange(language) %>%
  pull(language)

lang_batch1 <- langs_16[1:8]
lang_batch2 <- langs_16[9:16]

category_english_sample1 <- english_sample1 %>%
  filter(response %in% c("solid", "count noun", "shape")) %>%
  mutate(over_threshold = proportion >= threshold) %>%
  select(-block, -totcount, -count, -proportion, -response_numeric,
         -eng_response, -cleaned_response, -matching_response) %>%
  pivot_wider(names_from = "response", values_from = "over_threshold",
              values_fill = FALSE) %>%
  rename(count_noun = `count noun`)

ratings_aoa <- read_csv(here("data/predictors", "ratings_aoa.csv"), show_col_types = FALSE)
```

## Per-language unilemma counts and semantic categories

```{r supp-lang-n-tab, results="asis"}
lang_n_tbl <- english_sample1 %>%
  distinct(language, uni_lemma) %>%
  count(language, name = "N unilemmas") %>%
  arrange(language)

papaja::apa_table(lang_n_tbl,
                  caption = "Study 1A: Number of distinct unilemmas rated per Wordbank language list (English-speaking raters).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r"))
```

## Item-count reconciliation across studies

```{r supp-item-recon, results="asis"}
e2_supp <- read_csv(here("data/predictors", "english_sample2_props.csv"),
                    show_col_types = FALSE) %>% filter(!is.na(response))
k1_supp <- read_csv(here("data/predictors", "korean_sample1_props.csv"),
                    show_col_types = FALSE) %>%
  filter(!is.na(response), !uni_lemma %in% c("fish", "window"))
k2_supp <- read_csv(here("data/predictors", "korean_sample2_props.csv"),
                    show_col_types = FALSE) %>%
  filter(!is.na(response)) %>%
  mutate(Word = str_to_lower(Word)) %>%
  select(-uni_lemma) %>%
  filter(!Word %in% c("fish", "window"))
kor_map <- read_csv(here("data/ratings", "alllangs.csv"),
                    show_col_types = FALSE) %>%
  filter(language == "Korean", !is.na(uni_lemma)) %>%
  transmute(theword = item_definition, uni_lemma = str_to_lower(uni_lemma)) %>%
  distinct()
k2_matched <- merge(k2_supp, kor_map, by = "theword")

nb <- function(df, bl, col = "uni_lemma") df %>%
  filter(block == bl) %>% summarise(n = n_distinct(.data[[col]])) %>% pull(n)

recon_tbl <- dplyr::bind_rows(
  tibble::tibble(Sample = "Study 1A (English raters)",
                 Block = c("Solidity", "Count--mass", "Organizing feature"),
                 Items = c(nb(english_sample1, "solidity"),
                           nb(english_sample1, "count_mass"),
                           nb(english_sample1, "category_organization"))),
  tibble::tibble(Sample = "Study 1B (English raters)",
                 Block = "Organizing feature",
                 Items = nb(e2_supp, "category_organization", "word")),
  tibble::tibble(Sample = "Study 2A (Korean, in-country)",
                 Block = c("Solidity", "Count--mass", "Organizing feature"),
                 Items = c(nb(k2_supp, "solidity", "Word"),
                           nb(k2_supp, "count_mass", "Word"),
                           nb(k2_supp, "category_organization", "Word"))),
  tibble::tibble(Sample = "Study 2A, unilemma-matched",
                 Block = "All blocks",
                 Items = n_distinct(k2_matched$uni_lemma)),
  tibble::tibble(Sample = "Study 2B (Korean, online)",
                 Block = c("Solidity", "Count--mass", "Organizing feature"),
                 Items = c(nb(k1_supp, "solidity"),
                           nb(k1_supp, "count_mass"),
                           nb(k1_supp, "category_organization")))
)

papaja::apa_table(
  recon_tbl,
  caption = "Item counts by sample and block: distinct items with at least one retained rating.",
  note = paste0(
    "Study 1A blocks differ by a few nouns because not every noun retained ",
    "ratings in every block. The Study 1B word list lacked 40 of the 664 Study 1A nouns ",
    "(main text, Study 1B Materials). Study 2A items are English glosses of ",
    "the Korean words (two duplicate-gloss items excluded); the ",
    "unilemma-matched row counts the glosses matched to Korean CDI unilemmas ",
    "that enter cross-study analyses. Study 2B counts unilemmas after the ",
    "same duplicate-gloss exclusion. Pooled English counts in S7 exceed 664 ",
    "because the Study 1B and Study 3 lists contain nouns absent from Study 1A."
  ),
  placement = "!htb",
  escape = FALSE,
  align = c("l", "l", "r")
)
```

```{r supp-category-heatmap, fig.width = 7.2, fig.height = 5.4, fig.cap = "Semantic-category counts per language for Study 1A items (Wordbank semantic coding), for all 16 rated languages. Cell values are unilemma counts."}
# Drawn from the full Study 1A sampling frame (alllangs.csv) rather than the
# AoA analysis set, so Kiswahili and Spanish (Peruvian), which lack Wordbank
# AoA estimates, are included.
cat_counts <- read_csv(here("data/ratings", "alllangs.csv"),
                       show_col_types = FALSE) %>%
  filter(!is.na(uni_lemma)) %>%
  distinct(language, uni_lemma, category) %>%
  count(language, category, name = "n")

ggplot(cat_counts, aes(x = language, y = category, fill = n)) +
  geom_tile(color = "white", linewidth = 0.15) +
  geom_text(aes(label = n), size = 2.3, color = "grey15") +
  scale_fill_gradient(low = "#f7f7f7", high = "#4a6fa5") +
  labs(x = NULL, y = NULL, fill = "Count") +
  theme_minimal(base_size = 9) +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1, size = 6),
    axis.text.y = element_text(size = 7),
    legend.position = "right",
    panel.grid = element_blank()
  )
```

## Pairwise overlap of language-specific unilemma sets

```{r supp-lang-overlap-heatmap, fig.width = 7, fig.height = 6.2, fig.cap = "Pairwise Jaccard similarity between Study 1A noun lists (distinct unilemmas per language). Diagonal $= 1$."}
lemma_sets_df <- english_sample1 %>%
  distinct(language, uni_lemma) %>%
  group_by(language) %>%
  summarise(lemmas = list(unique(uni_lemma)), .groups = "drop")
lemma_sets <- lemma_sets_df$lemmas
names(lemma_sets) <- lemma_sets_df$language

langs_ord <- sort(names(lemma_sets))
jac_mat <- outer(langs_ord, langs_ord,
                 Vectorize(function(a, b) {
                   ia <- lemma_sets[[a]]
                   ib <- lemma_sets[[b]]
                   length(intersect(ia, ib)) / length(union(ia, ib))
                 }))
dimnames(jac_mat) <- list(langs_ord, langs_ord)

jac_long <- jac_mat %>%
  as.data.frame() %>%
  tibble::rownames_to_column("lang_a") %>%
  tidyr::pivot_longer(-lang_a, names_to = "lang_b", values_to = "jaccard")

ggplot(jac_long, aes(x = lang_a, y = lang_b, fill = jaccard)) +
  geom_tile(color = "white", linewidth = 0.08) +
  scale_fill_gradient2(low = "white", mid = "#c7d4e8", high = "#1c3d5a",
                       midpoint = 0.5, limits = c(0, 1)) +
  labs(x = NULL, y = NULL, fill = expression(Jaccard)) +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 5.5),
    axis.text.y = element_text(size = 5.5),
    panel.grid = element_blank()
  )
```

## Participant demographics (Study 1A)

```{r supp-demo-1a}
demo_paths <- list.files(here("data/demographics/first-sample-english"),
                         pattern = "\\.csv$", full.names = TRUE)
stopifnot(length(demo_paths) > 0)

english_demo_1a_all <- purrr::map_dfr(demo_paths, function(p) {
  read_csv(p, show_col_types = FALSE) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ suppressWarnings(as.character(.x))))
}) %>%
  janitor::clean_names()

subjects_1a <- read_csv(here("data/predictors", "english_sample1_raw.csv"),
                        show_col_types = FALSE) %>%
  filter(!is.na(subject)) %>%
  distinct(subject) %>%
  pull(subject)

n_export_rows <- nrow(english_demo_1a_all)
n_approved    <- sum(english_demo_1a_all$status == "APPROVED", na.rm = TRUE)

english_demo_1a <- english_demo_1a_all %>%
  filter(participant_id %in% subjects_1a) %>%
  distinct(participant_id, .keep_all = TRUE) %>%
  mutate(age = suppressWarnings(as.numeric(age)))

demo_sum <- tibble::tibble(
  Statistic = c("Platform submissions (all statuses)",
                "Approved submissions",
                "Participants matched to analyzed rating data",
                "Mean age (years)", "SD age", "Min age", "Max age"),
  Value = c(
    as.character(n_export_rows),
    as.character(n_approved),
    as.character(nrow(english_demo_1a)),
    sprintf("%.2f", mean(english_demo_1a$age, na.rm = TRUE)),
    sprintf("%.2f", sd(english_demo_1a$age,   na.rm = TRUE)),
    sprintf("%.0f", min(english_demo_1a$age, na.rm = TRUE)),
    sprintf("%.0f", max(english_demo_1a$age, na.rm = TRUE))
  )
)

demo_sex <- english_demo_1a %>%
  count(sex, name = "n") %>%
  mutate(pct = sprintf("%.1f%%", 100 * n / sum(n)))
```

```{r supp-demo-1a-tab1, results="asis"}
papaja::apa_table(demo_sum,
                  caption = "Study 1A demographics for participants matched to the analyzed rating data. Platform exports also include returned and timed-out submissions with no analyzed trials. The main-text N (336) counts distinct recorded identifiers in the rating data; one identifier has no matching demographic record, so demographic tables sum to 335. A session-logging error additionally left the identifier missing on a subset of retained trials (main text, Study 1A Participants).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r"))
```

```{r supp-demo-1a-tab2, results="asis"}
papaja::apa_table(demo_sex,
                  caption = "Study 1A: Self-reported sex (analyzed participants). One matched participant has demographic fields marked as consent-revoked in the platform export and contributes no age information.",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r", "r"))
```

## Distributions of rating totals and response proportions

```{r supp-rating-distributions, fig.width = 7.2, fig.height = 4.8, fig.cap = "\\textbf{Left:} Per-item rating totals by experimental block. \\textbf{Right:} Aggregated item-level proportions for solidity, count-noun, and shape responses."}
word_tot <- english_sample1 %>%
  distinct(language, uni_lemma, block, totcount)

p_tot <- ggplot(word_tot, aes(x = totcount)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "#4a6fa5", color = "white", linewidth = 0.2, alpha = 0.85) +
  facet_wrap(~block, ncol = 1, scales = "free_y") +
  labs(x = "Number of raters per item", y = "Density") +
  theme_minimal(base_size = 10)

resp_dim <- english_sample1 %>%
  dplyr::filter(matching_response %in% c("solid", "count", "shape"))

p_prop <- ggplot(resp_dim, aes(x = proportion)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30,
                 fill = "#6b8f71", color = "white", linewidth = 0.2, alpha = 0.85) +
  facet_wrap(~matching_response, ncol = 1, scales = "free_y") +
  labs(x = "Aggregated proportion (per unilemma by language)", y = "Density") +
  theme_minimal(base_size = 10)

gridExtra::grid.arrange(p_tot, p_prop, ncol = 2)
```

## Missing participant identifiers: sensitivity analysis

```{r missing-id-sensitivity}
mid_sens <- read_csv(here("data/predictors", "missing_id_sensitivity.csv"),
                     show_col_types = FALSE)
mid_shape <- mid_sens %>% filter(stat == "shape")
mid_resp <- read_csv(here("data/predictors", "missing_id_response_shares.csv"),
                     show_col_types = FALSE) %>%
  filter(block == "category_organization", response == "shape")
mid_share_na <- mid_resp %>% filter(id_status == "missing_id") %>% pull(share)
mid_share_id <- mid_resp %>% filter(id_status == "with_id") %>% pull(share)
```

A session-logging error left the participant identifier missing on 13.8% of retained Study 1A trials (main text, Study 1A Participants). All unattributed trials fall in the category-organization block (32.2% of that block); solidity and count–mass proportions are identical under either inclusion rule. Item-level mean shape is `r printnum(mid_shape$mean_with_id, digits = 2)` with identifier-present trials only versus `r printnum(mid_shape$mean_all, digits = 2)` with all trials (item-level $r$ = `r printnum(mid_shape$item_cor, digits = 2)`). Unattributed trials selected shape less often (`r printnum(100 * mid_share_na, digits = 0)`% vs. `r printnum(100 * mid_share_id, digits = 0)`%), so missingness may relate to responses. Under either rule, mean shape remains far below solidity and count-noun proportions (`r printnum(mid_sens$mean_all[mid_sens$stat == "solid"], digits = 2)` and `r printnum(mid_sens$mean_all[mid_sens$stat == "count"], digits = 2)`). Study 1A analyses are item-aggregated; trial-level models with participant random effects appear only in Study 3 on separate samples.

```{r missing-id-table}
apa_table(
  mid_sens %>%
    transmute(
      Statistic = dplyr::recode(stat, shape = "Shape", solid = "Solid",
                                count = "Count noun"),
      `Items` = n_items_all,
      `Mean (all trials)` = printnum(mean_all, digits = 3),
      `Mean (ID present)` = printnum(mean_with_id, digits = 3),
      `Item-level r` = printnum(item_cor, digits = 3),
      `Max |diff|` = printnum(max_abs_diff, digits = 3)
    ),
  caption = "Study 1A item-level proportions from all retained trials versus identifier-present trials only.",
  note = "Missing identifiers fall only in the category-organization block, so solidity and count-noun rows are identical by construction."
)
```

```{r missing-id-reruns}
# Identifier-present reruns of the downstream analyses that use the Study 1A
# shape measure: the per-language 80%-criterion shape counts and the primary
# Study 4 cross-linguistic model.
raw_1a_supp <- read_csv(here("data/predictors", "english_sample1_raw.csv"),
                        show_col_types = FALSE)
id_shape <- raw_1a_supp %>%
  filter(!is.na(subject), block == "category_organization") %>%
  group_by(word) %>%
  summarise(shape_id = mean(response == "shape"), .groups = "drop")

crit_all <- category_english_sample1 %>%
  group_by(language) %>% summarise(n = sum(shape), .groups = "drop")
crit_id <- english_sample1 %>%
  distinct(language, uni_lemma) %>%
  left_join(id_shape, by = c("uni_lemma" = "word")) %>%
  mutate(pass = !is.na(shape_id) & shape_id >= threshold) %>%
  group_by(language) %>% summarise(n = sum(pass), .groups = "drop")

# Same construction as the main-text primary model, with the Study 1A half
# of the averaged shape measure recomputed from identifier-present trials
# (nouns absent from Study 1B keep the Study 1A value alone, as in the
# main text).
s2_set <- unique(e2_supp$word)
mid_aoa_df <- ratings_aoa %>%
  left_join(id_shape, by = c("uni_lemma" = "word")) %>%
  mutate(mean_shape_id = ifelse(uni_lemma %in% s2_set,
                                (shape_id + shape_s2) / 2, shape_id)) %>%
  select(uni_lemma, aoa, logfreq, concreteness, solid, mean_shape_id,
         count_noun, category, language) %>%
  mutate(across(c(logfreq, concreteness, solid, mean_shape_id, count_noun),
                ~ as.numeric(scale(.x)))) %>%
  distinct()
mid_fit <- lmerTest::lmer(
  aoa ~ logfreq + concreteness + solid + mean_shape_id + count_noun +
        language + (1 | category) + (1 | uni_lemma),
  data = mid_aoa_df)
mid_cf <- coef(summary(mid_fit))["mean_shape_id", ]
```

Two further checks apply the identifier-present rule to the analyses downstream of the Study 1A shape measure. The primary Study 4 cross-linguistic model, refit with the identifier-present shape measure, gives $B$ = `r printnum(mid_cf[["Estimate"]], digits = 2)` ($SE$ = `r printnum(mid_cf[["Std. Error"]], digits = 2)`, $p$ = `r papaja::printp(mid_cf[["Pr(>|t|)"]])`), indistinguishable from the all-trials estimate in the main text. The per-language 80%-criterion shape sets grow under the identifier-present rule (up to `r max(crit_id$n)` nouns per language versus `r max(crit_all$n)` with all trials) because fewer raters per item make an 80% supermajority easier to reach; they remain far smaller than the corresponding solid and count sets.

# S3. Study 1A: Sixteen-language grids

```{r supp-correspondence-16a, fig.width=5.4, fig.height=3, fig.cap="(Part 1 of 2) Kernel density of per-word proportions for solidity, count, and shape, Study 1A (eight languages, alphabetical)."}
english_sample1 %>%
  filter(matching_response %in% c("solid", "count", "shape"),
         language %in% lang_batch1) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch1))) %>%
  ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language, ncol = 4) +
  labs(x = "Proportion of trials", y = "Density (words)") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 6),
    axis.text  = element_text(size = 5.5),
    strip.text = element_text(size = 5.5)
  )
```

```{r supp-correspondence-16b, fig.width=5.4, fig.height=3, fig.cap="(Part 2 of 2) Same as Part~1 for the remaining eight languages."}
english_sample1 %>%
  filter(matching_response %in% c("solid", "count", "shape"),
         language %in% lang_batch2) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch2))) %>%
  ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language, ncol = 4) +
  labs(x = "Proportion of trials", y = "Density (words)") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 6),
    axis.text  = element_text(size = 5.5),
    strip.text = element_text(size = 5.5)
  )
```

```{r supp-venn-16a, fig.width=5.4, fig.height=2.85, fig.cap="(Part 1 of 2) Area-proportional Euler diagrams (80\\% cutoff) for solid, count, and shape-organization, Study 1A (eight languages; same order as the density figure)."}
venns16a <- category_english_sample1 %>%
  filter(language %in% lang_batch1) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch1))) %>%
  arrange(language) %>%
  group_by(language) %>%
  nest() %>%
  mutate(plot = map2(data, language, ~ euler_panel(
    .x %>% select(`count noun` = count_noun, solid, shape), .y))) %>%
  pull(plot)

gridExtra::grid.arrange(grobs = venns16a, ncol = 4)
```

```{r supp-venn-16b, fig.width=5.4, fig.height=2.85, fig.cap="(Part 2 of 2) Area-proportional Euler diagrams for the remaining eight languages."}
venns16b <- category_english_sample1 %>%
  filter(language %in% lang_batch2) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch2))) %>%
  arrange(language) %>%
  group_by(language) %>%
  nest() %>%
  mutate(plot = map2(data, language, ~ euler_panel(
    .x %>% select(`count noun` = count_noun, solid, shape), .y))) %>%
  pull(plot)

gridExtra::grid.arrange(grobs = venns16b, ncol = 4)
```

# S4. Cross-sample shape correlations

```{r shape-4x4}
all_data_wide <- read_csv(here("data/predictors", "reliability_wide_alldata.csv"),
                          show_col_types = FALSE)

shape_wide <- all_data_wide %>%
  filter(response_norm == "shape") %>%
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(eng1, eng2, kor1, kor2) %>%
  rename(`Study 1A` = eng1, `Study 1B` = eng2,
         `Study 2B` = kor1, `Study 2A` = kor2)

cor4 <- cor(shape_wide, use = "pairwise.complete.obs")
cor4_df <- tibble::rownames_to_column(as.data.frame(round(cor4, 2)), "Sample")
```

```{r shape-4x4-table, results="asis"}
papaja::apa_table(cor4_df,
                  caption = "Pearson correlations of word-level shape-organization proportions across the four norming samples (Study 2A = in-country Korean; Study 2B = online Korean).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", rep("r", 4)))
```

# S5. Korean age-of-acquisition models

On the matched Korean Wordbank AoA item set, two models share the same rows and differ in which lexical-feature ratings enter. The English-predictor model includes Korean corpus log frequency, Brysbaert concreteness, and Study 1A/1B solidity, shape, and count-noun ratings. The Korean-predictor model includes the same Korean corpus log frequency and Korean-averaged solidity, shape, and count-noun ratings (Studies 2A and 2B), omitting concreteness because those norms are English-lemma ratings. Only a category random intercept is identifiable.

```{r korean-aoa-supp}
ratings_aoa <- read_csv(here("data/predictors", "ratings_aoa.csv"),
                        show_col_types = FALSE)
ratings_aoa_korean <- ratings_aoa %>% filter(language == "Korean")

kor_preds <- all_data_wide %>%
  select(uni_lemma, kor1, kor2, response_norm) %>%
  pivot_wider(names_from = response_norm, values_from = c(kor1, kor2),
              values_fill = 0) %>%
  mutate(
    mean_shape_kor = (kor1_shape + kor2_shape) / 2,
    mean_solid_kor = (kor1_solid + kor2_solid) / 2,
    mean_count_kor = (kor1_count + kor2_count) / 2
  ) %>%
  select(uni_lemma, mean_shape_kor, mean_solid_kor, mean_count_kor)

all_data_wide_kor_model <- ratings_aoa_korean %>%
  left_join(kor_preds, by = "uni_lemma") %>%
  # `freq` is already a within-language z-scored log relative frequency
  # (see analysis/build-ratings-aoa.R), so it is not logged again here.
  filter(!is.na(aoa), !is.na(freq)) %>%
  mutate(logfreq = freq) %>%
  drop_na(concreteness, solid, mean_shape, count_noun,
          mean_shape_kor, mean_solid_kor, mean_count_kor, category) %>%
  distinct(uni_lemma, .keep_all = TRUE) %>%
  mutate(across(
    c(logfreq, concreteness, solid, mean_shape, count_noun,
      mean_solid_kor, mean_shape_kor, mean_count_kor),
    ~ as.numeric(scale(.x))
  ))

kor_lmer_ctrl <- lmerControl(optimizer = "bobyqa",
                             optCtrl = list(maxfun = 2e5))

korean_eng_fit <- lmerTest::lmer(
  aoa ~ logfreq + concreteness + solid + mean_shape + count_noun +
        (1 | category),
  data = all_data_wide_kor_model, REML = FALSE, control = kor_lmer_ctrl)

korean_fit <- lmerTest::lmer(
  aoa ~ logfreq + mean_solid_kor + mean_shape_kor + mean_count_kor +
        (1 | category),
  data = all_data_wide_kor_model, REML = FALSE, control = kor_lmer_ctrl)

# Covariate-parity check: drop concreteness (an English-only norm) from the
# English-predictor model so the two models share the same covariate set.
korean_eng_fit_noconc <- lmerTest::lmer(
  aoa ~ logfreq + solid + mean_shape + count_noun + (1 | category),
  data = all_data_wide_kor_model, REML = FALSE, control = kor_lmer_ctrl)

coef_tbl <- function(fit) {
  k_tab <- coef(summary(fit))
  tibble::tibble(
    Predictor = rownames(k_tab),
    Estimate  = papaja::printnum(k_tab[, "Estimate"], digits = 2),
    SE        = papaja::printnum(k_tab[, "Std. Error"], digits = 2),
    df        = papaja::printnum(k_tab[, "df"], digits = 1),
    `t`       = papaja::printnum(k_tab[, "t value"], digits = 2),
    p         = papaja::printp(k_tab[, "Pr(>|t|)"])
  )
}
k_eng_out <- coef_tbl(korean_eng_fit)
k_out     <- coef_tbl(korean_fit)

korean_aic_supp <- c(
  eng = AIC(korean_eng_fit),
  kor = AIC(korean_fit)
)
korean_r2_supp <- if (requireNamespace("MuMIn", quietly = TRUE)) {
  c(
    eng = as.numeric(MuMIn::r.squaredGLMM(korean_eng_fit)[1, "R2m"]),
    kor = as.numeric(MuMIn::r.squaredGLMM(korean_fit)[1, "R2m"])
  )
} else {
  c(eng = NA_real_, kor = NA_real_)
}
```

```{r korean-aoa-table-eng, results="asis"}
papaja::apa_table(
  k_eng_out,
  caption = "Fixed effects for the English-predictor model of Korean age of acquisition (months).",
  note  = sprintf(
    "Random intercept for semantic category. Predictors are z-scored. N = %d. AIC = %s; marginal R^2 = %s.",
    nrow(all_data_wide_kor_model),
    papaja::printnum(korean_aic_supp["eng"], digits = 1),
    papaja::printnum(korean_r2_supp["eng"], digits = 2)
  ),
  placement = "!htb",
  escape = TRUE
)
```

```{r korean-aoa-table, results="asis"}
papaja::apa_table(
  k_out,
  caption = "Fixed effects for the Korean-predictor model of Korean age of acquisition (months).",
  note  = sprintf(
    "Random intercept for semantic category. Predictors are z-scored. N = %d. AIC = %s; marginal R^2 = %s. English-predictor model on the same items: AIC %s; marginal R^2 %s.",
    nrow(all_data_wide_kor_model),
    papaja::printnum(korean_aic_supp["kor"], digits = 1),
    papaja::printnum(korean_r2_supp["kor"], digits = 2),
    papaja::printnum(korean_aic_supp["eng"], digits = 1),
    papaja::printnum(korean_r2_supp["eng"], digits = 2)
  ),
  placement = "!htb",
  escape = TRUE
)
```

Because concreteness enters only the English-predictor model (the norms are English ratings), we also refit that model without it so the two models share the same covariate set; the comparison is unchanged (AIC `r papaja::printnum(AIC(korean_eng_fit_noconc), digits = 1)` vs. `r papaja::printnum(korean_aic_supp["kor"], digits = 1)` for the Korean-predictor model).

# S6. Study 2A: Strict vs. broad shape coding

Strict coding counts shape only when it was the sole selected feature; broad coding counts shape whenever it was selected alone or with another feature.

```{r kor2-supp, results="asis"}
kor2_sens <- read_csv(here("data/predictors", "study2a_shape_coding_sensitivity.csv"),
                      show_col_types = FALSE)
kor2_v <- function(s) kor2_sens$value[kor2_sens$statistic == s]

sens_tbl <- tibble::tibble(
  Description = c(
    "Organizing-block trials (weighted N)",
    "Shape only (strict)",
    "Shape + other feature (broad only)",
    "N items",
    "Mean P(shape), strict",
    "SD, strict",
    "Mean P(shape), broad",
    "SD, broad"
  ),
  Value = c(
    prettyNum(kor2_v("organizing_trials_total"), big.mark = ","),
    prettyNum(kor2_v("shape_only_trials"), big.mark = ","),
    prettyNum(kor2_v("shape_plus_other_trials"), big.mark = ","),
    as.character(kor2_v("n_items")),
    sprintf("%.3f", kor2_v("mean_strict")),
    sprintf("%.3f", kor2_v("sd_strict")),
    sprintf("%.3f", kor2_v("mean_broad")),
    sprintf("%.3f", kor2_v("sd_broad"))
  )
)

papaja::apa_table(sens_tbl,
                  caption = "Study 2A multiple-selection coding. Values match the main text. P(shape) = item-level shape proportion under each coding rule.",
                  placement = "!htb",
                  escape = FALSE,
                  align = c("l", "r"))
```

# S7. Pooled English rater counts

Organizing-feature ratings were pooled across English samples that shared a procedure, then re-scored under the 80% rule. Solidity and count–mass criteria remain from Study 1A (the only sample that collected those blocks).

```{r eng-agg-load}
eng_agg_raters <- read_csv(here("data/predictors", "english_aggregation_rater_summary.csv"),
                           show_col_types = FALSE)
eng_agg_crit <- read_csv(here("data/predictors", "english_aggregation_80pct_criterion.csv"),
                         show_col_types = FALSE)
eng_agg_sub <- read_csv(here("data/predictors", "english_aggregation_subsample_sensitivity.csv"),
                        show_col_types = FALSE)
eng_agg_ms <- read_csv(here("data/predictors", "english_aggregation_ms_blocked_criterion.csv"),
                       show_col_types = FALSE)

agg_rater_tbl <- eng_agg_raters %>%
  transmute(
    Procedure = procedure,
    Trials = prettyNum(n_trials, big.mark = ","),
    Words = n_unique_words,
    `Median raters/word` = med_raters_per_word,
    `Mean shape` = sprintf("%.2f", mean_shape)
  )

agg_crit_tbl <- eng_agg_crit %>%
  filter(language %in% c("English (American)", "Korean",
                         "Mandarin (Beijing)", "Japanese")) %>%
  transmute(
    Pool = dplyr::recode(pool,
                         `1A_only` = "Study 1A only",
                         `pooled_FC_random` = "Pooled FC random"),
    Language = language,
    `N shape` = n_shape,
    `N intersect` = n_intersect,
    `Median raters` = med_raters_shape,
    `Mean shape` = round(mean_shape, 2)
  )

agg_sub_tbl <- eng_agg_sub %>%
  transmute(
    `Raters per word` = n_raters,
    `Eligible words` = n_words,
    `Mean pass rate` = sprintf("%.1f\\%%", 100 * mean_pass_rate),
    `Words usually passing` = n_usually_pass
  )

agg_ms_tbl <- eng_agg_ms %>%
  transmute(
    Pool = dplyr::recode(pool,
                         ms_blocked_pilot = "MS grouped pilot",
                         pooled_ms_blocked = "Pooled MS grouped"),
    Words = n_words,
    `Words on 1A set` = n_words_on_1a_set,
    `Pct shape >= 80` = round(pct_shape_ge_80, 1),
    `Pct on 1A set` = round(pct_shape_ge_80_on_1a, 1),
    `Median raters` = med_raters,
    `Mean broad shape` = round(mean_broad_shape, 2)
  )
```

Pooling Studies 1A, 1B, and the forced-choice random cell of Study 3 raised the median number of organizing-feature raters per noun from `r eng_agg_raters %>% filter(procedure == "Study 1A (FC random)") %>% pull(med_raters_per_word)` to `r eng_agg_raters %>% filter(procedure == "Pooled FC random (1A + 1B + 2x2)") %>% pull(med_raters_per_word)`, left mean shape slightly lower, and changed 80% shape counts only marginally. Pooled word counts exceed the 664-noun Study 1A set because the Study 1B and Study 3 lists contain nouns absent from Study 1A; pooled rows count every word with at least one rating in any contributing sample.

```{r eng-agg-raters, results="asis"}
papaja::apa_table(
  agg_rater_tbl,
  caption = "Organizing-feature trials by procedure. FC = forced choice with random word order; MS blocked = multiple selection with semantic grouping.",
  placement = "!htb",
  escape = FALSE
)
```

```{r eng-agg-crit, results="asis"}
papaja::apa_table(
  agg_crit_tbl,
  caption = "80\\% cutoff counts in four languages: Study 1A shape ratings only versus pooled forced-choice random-order shape ratings. Solid and count criteria are from Study 1A in both rows.",
  placement = "!htb",
  escape = FALSE
)
```

On the pooled forced-choice trials, $n \in \{7, 9, 11, 13\}$ raters were redrawn per noun. Even at $n = 11$, fewer than 3% of items passed the 80% shape rule on average.

```{r eng-agg-sub, results="asis"}
papaja::apa_table(
  agg_sub_tbl,
  caption = "Subsample sensitivity on pooled forced-choice random-order trials. Mean pass rate = share of 400 resamples per word reaching $\\geq 80\\%$ shape; usually passing = pass rate $\\geq .5$.",
  placement = "!htb",
  escape = FALSE
)
```

Pooling the grouped multiple-selection pilot and the Study 3 grouped cell yields mean broad shape near the Study 3 cell estimates, but only `r eng_agg_ms %>% filter(pool == "pooled_ms_blocked") %>% pull(pct_shape_ge_80_on_1a) %>% round(1)`% of Study 1A nouns meet an 80% broad-shape cutoff, because most items in the $2 \times 2$ cells have few raters each.

```{r eng-agg-ms, results="asis"}
papaja::apa_table(
  agg_ms_tbl,
  caption = "Multiple-selection grouped samples (broad coding). The on-1A columns restrict to nouns that appear in Study 1A.",
  placement = "!htb",
  escape = FALSE
)
```

# S8. Leave-one-language-out refits of the Study 4 model

```{r lolo-load}
lolo_all <- read_csv(here("data/predictors", "aoa_lolo_sensitivity.csv"),
                     show_col_types = FALSE)
lolo_pool <- lolo_all %>% filter(dropped_language == "none (pooled model)")
lolo      <- lolo_all %>% filter(dropped_language != "none (pooled model)")

lolo_tbl_out <- lolo_all %>%
  transmute(
    `Language dropped` = dropped_language,
    `N obs` = prettyNum(n_obs, big.mark = ","),
    `B (shape)` = printnum(estimate, digits = 2),
    SE = printnum(se, digits = 2),
    `95\\% CI` = sprintf("[%s, %s]",
                         printnum(ci_low, digits = 2),
                         printnum(ci_high, digits = 2)),
    p = papaja::printp(p)
  )
```

The primary cross-linguistic model of Study 4 was refit `r nrow(lolo)` times, each time dropping all items contributed by one language. The shape coefficient stayed negative in every refit and varied between `r printnum(min(lolo$estimate), digits = 2)` and `r printnum(max(lolo$estimate), digits = 2)`, against a pooled estimate of `r printnum(lolo_pool$estimate, digits = 2)`, and the Wald test fell below .05 in `r sum(lolo$p < .05)` of the `r nrow(lolo)` refits. No single language's item set is necessary for the association, although this check retains nearly all unilemmas on each pass (the shape predictor is unilemma-level), so it speaks to language-specific item sets rather than to the unilemma-level predictor; the more stringent leave-one-category-out refits appear in the main text (Study 4 Results).

```{r lolo-table, results="asis"}
papaja::apa_table(
  lolo_tbl_out,
  caption = "Leave-one-language-out refits of the Study 4 cross-linguistic model. Each row drops one language's item-by-language observations and reports the shape coefficient from the refit (months of age of acquisition per SD of shape rating).",
  note = "First row is the pooled model reported in the main text. Wald 95\\% confidence intervals. All other model terms, including the language fixed effect and the crossed category and unilemma random intercepts, are as in the main model.",
  placement = "!htb",
  escape = FALSE,
  align = c("l", "r", "r", "r", "r", "r")
)
```

# S9. Bayesian refit and power sensitivity for the shape--AoA coefficient

```{r bayes-power-load}
bp_bayes <- read_csv(here("data/predictors", "aoa_bayes_summary.csv"),
                     show_col_types = FALSE)
bp_power <- read_csv(here("data/predictors", "aoa_power_sensitivity.csv"),
                     show_col_types = FALSE)

bayes_tbl <- bp_bayes %>%
  transmute(
    Model = model,
    `Post. mean` = printnum(post_mean, digits = 2),
    `Post. SD` = printnum(post_sd, digits = 2),
    `95\\% CrI` = sprintf("[%s, %s]",
                          printnum(ci_low, digits = 2),
                          printnum(ci_high, digits = 2)),
    `$P(B < 0)$` = printnum(p_negative, digits = 3),
    `\\% in ROPE` = printnum(100 * pct_in_rope, digits = 1),
    `\\% below ROPE` = printnum(100 * pct_below_rope, digits = 1),
    `Max $\\hat{R}$` = printnum(max_rhat_b, digits = 3)
  )

power_tbl <- bp_power %>%
  filter(design == "English (American) only") %>%
  transmute(`True shape slope` = printnum(true_shape_slope, digits = 2),
            Power = printnum(power, digits = 2))
pw_n_eng <- bp_power %>% filter(design == "English (American) only") %>%
  slice(1) %>% pull(n_items)

bp_primary <- bp_bayes %>% filter(model == "cross-linguistic (primary)")
```

Both models were refit in `brms` with weakly informative normal(0, 5) priors on the regression coefficients, four chains of 4,000 iterations (1,000 warmup), and `adapt_delta` = 0.95. The region of practical equivalence (ROPE) is $\pm 0.1$ SD of age of acquisition, or $\pm$ `r printnum(bp_primary$rope_high, digits = 2)` months. For the primary cross-linguistic model, `r printnum(100 * bp_primary$p_negative, digits = 1)`% of the posterior mass for the shape coefficient lies below zero and `r printnum(100 * bp_primary$pct_below_rope, digits = 1)`% lies below the ROPE. The direction of the effect is therefore well supported while its practical size is not: about `r printnum(100 * bp_primary$pct_in_rope, digits = 0)`% of the posterior is compatible with a negligible effect.

```{r bayes-table, results="asis"}
papaja::apa_table(
  bayes_tbl,
  caption = "Posterior summaries for the shape coefficient in the Bayesian refits of the Study 4 models.",
  note = "Coefficients are in months of age of acquisition per SD of shape rating. CrI = credible interval; ROPE = region of practical equivalence, $\\pm 0.1$ SD of age of acquisition.",
  placement = "!htb",
  escape = FALSE,
  align = c("l", rep("r", 7))
)
```

The single-language design cannot arbitrate the same question. Simulating age of acquisition from the English-only design ($N$ = `r pw_n_eng`) at a range of true slopes, with the fitted category-intercept and residual variances, gives the rejection rates below.

```{r power-table, results="asis"}
papaja::apa_table(
  power_tbl,
  caption = sprintf("Simulated power of the English-only Study 4 design (N = %d items, 500 simulations per row) to detect a shape effect of a given size.", pw_n_eng),
  note = sprintf(
    "Slopes are in months per SD of shape rating; the pooled cross-linguistic estimate is %s. The analytic minimal detectable effect at 80\\%% power is %s for the English-only design and %s for the Korean-predictor design.",
    printnum(lolo_pool$estimate, digits = 2),
    printnum(abs(bp_power$true_shape_slope[bp_power$design == "English-only analytic MDE (80% power)"]), digits = 2),
    printnum(abs(bp_power$true_shape_slope[bp_power$design == "Korean-predictor analytic MDE (80% power)"]), digits = 2)),
  placement = "!htb",
  escape = FALSE,
  align = c("r", "r")
)
```
