---
title: "Supplementary materials: Lexical statistics in early noun vocabularies (shape bias project)"
author: "Samah Abdelrahim & Michael C. Frank (authorship as in main text)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: pdflatex
    toc: false
header-includes:
  - \usepackage{graphicx}
  - \usepackage{booktabs}
  - \usepackage{threeparttable}
  - \usepackage{float}
  - \usepackage{etoolbox}
  - \AtBeginEnvironment{table}{\small\setlength{\tabcolsep}{4.5pt}}
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.path   = "data/figures/supplement-",
  fig.align  = "center",
  dev        = "pdf",
  dpi        = 300,
  out.width  = "0.98\\textwidth",
  fig.width  = 5.4,
  fig.height = 3.0
)

library(here)
library(tidyverse)
library(papaja)
library(psych)
library(ggvenn)
library(gridExtra)
library(janitor)
library(lmerTest)

source(here("helper.R"))
```

This document supplements the main manuscript *Lexical statistics and the shape bias* (same data and code repository). It provides (S1) kernel-density plots of solidity, count, and shape ratings for **all 16** unilemma languages in Study~1A, **split into two page-sized figures**; (S2) 80% cutoff Venn diagrams, likewise split; (S3) the $4 \times 4$ correlation matrix of **word-level shape proportions** across English samples 1--2 and Korean samples 1--2; (S4) the Korean-only age-of-acquisition model reported as a robustness check in the main text; (S5) numeric details of the Study~2B strict- vs. broad-shape coding sensitivity analysis; and (S6) pointers to **browser-ready jsPsych implementations** of the rating tasks (Studies~1A and 2A), described below. **Study~1A Materials coverage** (referenced in the main manuscript): **per-language unilemma counts** and a **semantic-category composition** table and heatmap (linked via `ratings_aoa.csv`); **pairwise Jaccard overlap** between the 16 language-specific noun lists; **participant demographics** from Prolific CSV exports; and **distributions** of per-item rating totals (`totcount`) and response proportions across blocks.

# Experiment implementations (Studies 1A and 2A)

The repository folder `experiment files/` (at the project root) contains the exact **HTML + JavaScript** tasks used for online collection. They mirror the procedures described in the main text:

- **Study 1A (English, 16-language unilemma lists):** Open `solidity-ratings.html`, which loads `solidity-ratings.js` (embedded word lists and jsPsych timeline). Participants completed three blocks—solidity, count–mass syntax, and category-organizing feature—in random order (20 nouns per block). On test trials, each *uni_lemma* appears alone as the prompt for solidity and count–mass; the organizing-feature trial asks whether the lemma *belongs to a category that is organized by:* with responses *shape* / *color* / *material* / *none of these*.
- **Study 2A (Korean):** Open `vocab_comp.html`, which loads `words.js` (defines global `words_array`) then `vocab_comp.js`. Instructions follow the clarified wording used in Study 2A; test-trial structure matches Study 1A for the corresponding blocks.

Supporting assets include local copies of jsPsych plugins under `experiment files/jspsych/` and shared styling (`solidity.css`). Opening the HTML files directly from disk (`file://`) works with the scripts as checked in (classic scripts; no ES-module imports required). Serving the folder over a local HTTP server is optional for deployment or debugging.

# Study 1A: Lexicon composition, cross-language overlap, demographics, and rating distributions

```{r load-eng1}
english_sample1 <- read_csv(here("data/predictors", "english_sample1_props.csv"),
                           show_col_types = FALSE) %>%
  filter(!is.na(response)) %>%
  distinct()
threshold <- 0.8

langs_16 <- english_sample1 %>%
  distinct(language) %>%
  arrange(language) %>%
  pull(language)

lang_batch1 <- langs_16[1:8]
lang_batch2 <- langs_16[9:16]

category_english_sample1 <- english_sample1 %>%
  filter(response %in% c("solid", "count noun", "shape")) %>%
  mutate(over_threshold = proportion >= threshold) %>%
  select(-block, -totcount, -count, -proportion, -response_numeric,
         -eng_response, -cleaned_response, -matching_response) %>%
  pivot_wider(names_from = "response", values_from = "over_threshold",
              values_fill = FALSE) %>%
  rename(count_noun = `count noun`)

ratings_aoa <- read_csv(here("data/predictors", "ratings_aoa.csv"), show_col_types = FALSE)
```

## Per-language unilemma counts and semantic categories

```{r supp-lang-n-tab, results="asis"}
lang_n_tbl <- english_sample1 %>%
  distinct(language, uni_lemma) %>%
  count(language, name = "N unilemmas") %>%
  arrange(language)

papaja::apa_table(lang_n_tbl,
                  caption = "Study~1A: Number of distinct unilemmas rated per Wordbank language list (English-speaking raters).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r"))
```

```{r supp-category-heatmap, fig.width = 7.2, fig.height = 5.4, fig.cap = "Semantic category counts per language for Study~1A items (categories taken from \\texttt{ratings\\_aoa.csv}, aligned with AoA regression syntax). Cell value is the number of unilemmas in each category--language cell."}
cat_counts <- ratings_aoa %>%
  distinct(language, uni_lemma, category) %>%
  count(language, category, name = "n")

ggplot(cat_counts, aes(x = language, y = category, fill = n)) +
  geom_tile(color = "white", linewidth = 0.15) +
  geom_text(aes(label = n), size = 2.3, color = "grey15") +
  scale_fill_gradient(low = "#f7f7f7", high = "#8C1515") +
  labs(x = NULL, y = NULL, fill = "Count") +
  theme_minimal(base_size = 9) +
  theme(
    axis.text.x = element_text(angle = 55, hjust = 1, size = 6),
    axis.text.y = element_text(size = 7),
    legend.position = "right",
    panel.grid = element_blank()
  )
```

```{r supp-category-wide-tab, results="asis"}
cat_wide <- cat_counts %>%
  pivot_wider(names_from = language, values_from = n, values_fill = 0)

papaja::apa_table(cat_wide,
                  caption = "Study~1A: Same semantic-category counts as the heatmap (wide format). Zeros indicate no noun from that category in that language list.",
                  note  = "Categories follow Wordbank semantic coding linked in predictor construction.",
                  placement = "!htb",
                  escape = TRUE)
```

## Pairwise overlap of language-specific unilemma sets (Jaccard index)

```{r supp-lang-overlap-heatmap, fig.width = 7, fig.height = 6.2, fig.cap = "Pairwise Jaccard similarity $|L_i \\cap L_j| / |L_i \\cup L_j|$ between Study~1A noun lists $L_i$, defined over distinct \\texttt{uni\\_lemma} values per \\texttt{language} in \\texttt{english\\_sample1\\_props.csv}. Diagonal $= 1$."}
lemma_sets_df <- english_sample1 %>%
  distinct(language, uni_lemma) %>%
  group_by(language) %>%
  summarise(lemmas = list(unique(uni_lemma)), .groups = "drop")
lemma_sets <- lemma_sets_df$lemmas
names(lemma_sets) <- lemma_sets_df$language

langs_ord <- sort(names(lemma_sets))
jac_mat <- outer(langs_ord, langs_ord,
                 Vectorize(function(a, b) {
                   ia <- lemma_sets[[a]]
                   ib <- lemma_sets[[b]]
                   length(intersect(ia, ib)) / length(union(ia, ib))
                 }))
dimnames(jac_mat) <- list(langs_ord, langs_ord)

jac_long <- jac_mat %>%
  as.data.frame() %>%
  tibble::rownames_to_column("lang_a") %>%
  tidyr::pivot_longer(-lang_a, names_to = "lang_b", values_to = "jaccard")

ggplot(jac_long, aes(x = lang_a, y = lang_b, fill = jaccard)) +
  geom_tile(color = "white", linewidth = 0.08) +
  scale_fill_gradient2(low = "white", mid = "#c7d4e8", high = "#1c3d5a", midpoint = 0.5,
                       limits = c(0, 1)) +
  labs(x = NULL, y = NULL, fill = expression(Jaccard)) +
  theme_minimal(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 5.5),
    axis.text.y = element_text(size = 5.5),
    panel.grid = element_blank()
  )
```

## Participant demographics (Study 1A; Prolific)

```{r supp-demo-1a}
demo_paths <- list.files(here("data/demographics/first-sample-english"),
                         pattern = "\\.csv$", full.names = TRUE)
stopifnot(length(demo_paths) > 0)

english_demo_1a <- purrr::map_dfr(demo_paths, function(p) {
  read_csv(p, show_col_types = FALSE) %>%
    dplyr::mutate(dplyr::across(dplyr::everything(), ~ suppressWarnings(as.character(.x))))
}) %>%
  janitor::clean_names()

english_demo_1a <- english_demo_1a %>%
  mutate(age = suppressWarnings(as.numeric(age)))

demo_sum <- tibble::tibble(
  Statistic = c("N participants (rows in combined Prolific exports)",
                 "Mean age (years)", "SD age", "Min age", "Max age"),
  Value = c(
    as.character(nrow(english_demo_1a)),
    sprintf("%.2f", mean(english_demo_1a$age, na.rm = TRUE)),
    sprintf("%.2f", sd(english_demo_1a$age,   na.rm = TRUE)),
    sprintf("%.0f", min(english_demo_1a$age, na.rm = TRUE)),
    sprintf("%.0f", max(english_demo_1a$age, na.rm = TRUE))
  )
)

demo_sex <- english_demo_1a %>%
  count(sex, name = "n") %>%
  mutate(pct = sprintf("%.1f%%", 100 * n / sum(n)))
```

```{r supp-demo-1a-tab1, results="asis"}
papaja::apa_table(demo_sum,
                  caption = "Study~1A: Demographic summary from merged Prolific export files (\\texttt{data/demographics/first-sample-english/}).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r"))
```

```{r supp-demo-1a-tab2, results="asis"}
papaja::apa_table(demo_sex,
                  caption = "Study~1A: Self-reported sex (Prolific).",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r", "r"))
```

## Distributions of rating totals and response proportions (Study 1A)

```{r supp-rating-distributions, fig.width = 7.2, fig.height = 4.8, fig.cap = "\\textbf{Left:} Distribution of per-item--language totals (number of participant judgments contributing to each aggregated proportion), by experimental block. \\textbf{Right:} Distribution of aggregated proportions for solidity, count-noun, and shape responses (coded dimension of proportion column)."}
word_tot <- english_sample1 %>%
  distinct(language, uni_lemma, block, totcount)

p_tot <- ggplot(word_tot, aes(x = totcount)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "#8C1515", color = "white", linewidth = 0.2, alpha = 0.85) +
  facet_wrap(~block, ncol = 1, scales = "free_y") +
  labs(x = "totcount (raters contributing to item)", y = "Density") +
  theme_minimal(base_size = 10)

resp_dim <- english_sample1 %>%
  dplyr::filter(matching_response %in% c("solid", "count", "shape"))

p_prop <- ggplot(resp_dim, aes(x = proportion)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30,
                 fill = "#4a6fa5", color = "white", linewidth = 0.2, alpha = 0.85) +
  facet_wrap(~matching_response, ncol = 1, scales = "free_y") +
  labs(x = "Aggregated proportion (per uni_lemma × language)", y = "Density") +
  theme_minimal(base_size = 10)

gridExtra::grid.arrange(p_tot, p_prop, ncol = 2)
```

# Study 1A: Sixteen languages (full grids, two panels per plot type)

<!-- `english_sample1`, `langs_16`, batches, Venns prepped in chunk `load-eng1` above. -->

```{r supp-correspondence-16a, fig.width=5.4, fig.height=3, fig.cap="(Part 1 of 2) Kernel density of per-word proportions for solidity, count, and shape responses, Study~1A (eight unilemma languages, alphabetical a--h). See Part~2 for remaining languages."}
english_sample1 %>%
  filter(matching_response %in% c("solid", "count", "shape"),
         language %in% lang_batch1) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch1))) %>%
  ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language, ncol = 4) +
  labs(x = "Proportion of trials", y = "Density (words)") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 6),
    axis.text  = element_text(size = 5.5),
    strip.text = element_text(size = 5.5)
  )
```

```{r supp-correspondence-16b, fig.width=5.4, fig.height=3, fig.cap="(Part 2 of 2) Same as Part~1 for the remaining eight unilemma languages (alphabetical i--p)."}
english_sample1 %>%
  filter(matching_response %in% c("solid", "count", "shape"),
         language %in% lang_batch2) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch2))) %>%
  ggplot(aes(x = proportion, color = response, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ language, ncol = 4) +
  labs(x = "Proportion of trials", y = "Density (words)") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 6),
    axis.text  = element_text(size = 5.5),
    strip.text = element_text(size = 5.5)
  )
```

```{r supp-venn-16a, fig.width=5.4, fig.height=2.85, fig.cap="(Part 1 of 2) Venn diagrams (80\\% cutoff) for solid, count, and shape-organization, Study~1A (eight languages; same ordering as the density figure)."}
venns16a <- category_english_sample1 %>%
  filter(language %in% lang_batch1) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch1))) %>%
  group_by(language) %>%
  nest() %>%
  mutate(plot = map2(data, language, ~ ggvenn(
    .x, c("count_noun", "solid", "shape"),
    show_percentage = FALSE, text_size = 1.45, set_name_size = 1.35) +
    labs(y = .y) +
    theme(axis.title.y = element_text(angle = 90, size = 4, vjust = 0.5),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          axis.line.y = element_blank()))) %>%
  pull(plot)

gridExtra::grid.arrange(grobs = venns16a, ncol = 4)
```

```{r supp-venn-16b, fig.width=5.4, fig.height=2.85, fig.cap="(Part 2 of 2) Venn diagrams for the remaining eight languages."}
venns16b <- category_english_sample1 %>%
  filter(language %in% lang_batch2) %>%
  mutate(language = forcats::fct_relevel(factor(language), as.character(lang_batch2))) %>%
  group_by(language) %>%
  nest() %>%
  mutate(plot = map2(data, language, ~ ggvenn(
    .x, c("count_noun", "solid", "shape"),
    show_percentage = FALSE, text_size = 1.45, set_name_size = 1.35) +
    labs(y = .y) +
    theme(axis.title.y = element_text(angle = 90, size = 4, vjust = 0.5),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          axis.line.y = element_blank()))) %>%
  pull(plot)

gridExtra::grid.arrange(grobs = venns16b, ncol = 4)
```

# Cross-sample shape correlations

```{r shape-4x4}
all_data_wide <- read_csv(here("data/predictors", "reliability_wide_alldata.csv"),
                          show_col_types = FALSE)

shape_wide <- all_data_wide %>%
  filter(response_norm == "shape") %>%
  distinct(uni_lemma, .keep_all = TRUE) %>%
  select(eng1, eng2, kor1, kor2)

cor4 <- cor(shape_wide, use = "pairwise.complete.obs")
cor4_df <- tibble::rownames_to_column(as.data.frame(cor4), "Sample")
```

```{r shape-4x4-table, results="asis"}
papaja::apa_table(cor4_df,
                  caption = "Pearson correlations of word-level shape-organization proportions across the four norming samples (Study 1A = eng1, Study 1B = eng2, Study 2A = kor1, Study 2B = kor2).",
                  note  = "Lower triangle matches psych::corr.test output; diagonal is 1.",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", rep("r", 4)))
```

# Korean-only age-of-acquisition model

```{r korean-aoa-supp}
ratings_aoa <- read_csv(here("data/predictors", "ratings_aoa.csv"),
                        show_col_types = FALSE)
ratings_aoa_korean <- ratings_aoa %>% filter(language == "Korean")

all_data_wide_kor_model <- all_data_wide %>%
  select(uni_lemma, kor1, kor2, response_norm) %>%
  pivot_wider(names_from = response_norm, values_from = c(kor1, kor2),
              values_fill = 0) %>%
  mutate(
    mean_shape_kor = (kor1_shape + kor2_shape) / 2,
    mean_solid_kor = (kor1_solid + kor2_solid) / 2,
    mean_count_kor = (kor1_count + kor2_count) / 2
  ) %>%
  left_join(ratings_aoa_korean, by = "uni_lemma") %>%
  select(-ends_with(".y")) %>%
  rename_with(~ str_remove_all(., "\\.x$"), everything()) %>%
  filter(!is.na(aoa)) %>%
  distinct() %>%
  janitor::clean_names() %>%
  drop_na() %>%
  filter(!is.na(freq) & freq > 0) %>%
  mutate(logfreq = log(freq)) %>%
  drop_na(concreteness, solid, mean_shape, count_noun) %>%
  distinct()

korean_fit <- lmerTest::lmer(
  aoa ~ logfreq + concreteness + mean_solid_kor + mean_shape_kor + mean_count_kor +
        (1 | category),
  data = all_data_wide_kor_model
)
k_tab <- coef(summary(korean_fit))
k_out <- tibble::tibble(
  Predictor = rownames(k_tab),
  Estimate    = papaja::printnum(k_tab[, "Estimate"], digits = 2),
  SE          = papaja::printnum(k_tab[, "Std. Error"], digits = 2),
  df          = papaja::printnum(k_tab[, "df"], digits = 1),
  `t`         = papaja::printnum(k_tab[, "t value"], digits = 2),
  p           = papaja::printp(k_tab[, "Pr(>|t|)"])
)
```

```{r korean-aoa-table, results="asis"}
papaja::apa_table(k_out,
                  caption = "Fixed effects for the Korean-only model predicting age of acquisition (months) from corpus log-frequency, concreteness, and Korean-averaged solidity, shape, and count-noun ratings.",
                  note  = "Random intercept for semantic category. Predictors use the same scaling as in the main text where applicable.",
                  placement = "!htb",
                  escape = TRUE)
```

# Study 2B: Strict vs. broad shape coding

The main manuscript reports counts and means for strict (``shape'' as the only selected feature) versus broad (``shape'' selected alone or with another feature) coding. Those values are computed in `manuscript.Rmd` chunk `kor2-sensitivity` from `data/predictors/korean_sample2_props.csv`. For submission, paste the latest inline values from the rendered main text here if the journal requires a static supplement table.

```{r kor2-supp, results="asis"}
korean_sample2 <- read_csv(here("data/predictors", "korean_sample2_props.csv"),
                           show_col_types = FALSE) %>%
  mutate(Word = stringr::str_to_lower(Word)) %>%
  filter(!is.na(response)) %>%
  distinct() %>%
  filter(!Word %in% c("fish", "window"))

shape_strict_kor2 <- korean_sample2 %>%
  filter(block == "category_organization", matching_response == "shape") %>%
  distinct(Word, proportion) %>%
  rename(prop_strict = proportion)

shape_broad_kor2 <- korean_sample2 %>%
  filter(block == "category_organization") %>%
  filter(matching_response == "shape" |
         (matching_response == "multiple" &
            stringr::str_detect(cleaned_response, "form"))) %>%
  group_by(Word) %>%
  summarise(prop_broad = sum(proportion), .groups = "drop")

shape_kor2_compare <- dplyr::full_join(shape_strict_kor2, shape_broad_kor2,
                                       by = "Word") %>%
  mutate(
    prop_strict = ifelse(is.na(prop_strict), 0, prop_strict),
    prop_broad  = ifelse(is.na(prop_broad),  prop_strict, prop_broad)
  )

n_trials_total <- korean_sample2 %>%
  filter(block == "category_organization") %>%
  distinct(Word, totcount) %>%
  summarise(n = sum(totcount, na.rm = TRUE)) %>% pull(n)

n_shape_only <- korean_sample2 %>%
  filter(block == "category_organization", matching_response == "shape") %>%
  summarise(n = sum(proportion * totcount, na.rm = TRUE)) %>% pull(n) %>% round()

n_shape_multi <- korean_sample2 %>%
  filter(block == "category_organization",
         matching_response == "multiple",
         stringr::str_detect(cleaned_response, "form")) %>%
  summarise(n = sum(proportion * totcount, na.rm = TRUE)) %>% pull(n) %>% round()

sens_tbl <- tibble::tibble(
  Description = c(
    "Organizing block trials (total, weighted N)",
    "``Shape only'' (strict)",
    "Shape + other feature (broad)",
    "Mean P(shape), strict",
    "SD, strict",
    "Mean P(shape), broad",
    "SD, broad"
  ),
  Value = c(
    as.character(prettyNum(n_trials_total, big.mark = ",")),
    as.character(prettyNum(n_shape_only, big.mark = ",")),
    as.character(prettyNum(n_shape_multi, big.mark = ",")),
    sprintf("%.3f", mean(shape_kor2_compare$prop_strict, na.rm = TRUE)),
    sprintf("%.3f", sd(shape_kor2_compare$prop_strict,   na.rm = TRUE)),
    sprintf("%.3f", mean(shape_kor2_compare$prop_broad,  na.rm = TRUE)),
    sprintf("%.3f", sd(shape_kor2_compare$prop_broad,    na.rm = TRUE))
  )
)

papaja::apa_table(sens_tbl,
                  caption = "Sensitivity analysis for Study~2B multiple-selection trials (per main text). P(shape) = item-level shape proportion under each coding rule.",
                  placement = "!htb",
                  escape = TRUE,
                  align = c("l", "r"))
```
