---
title: "lexical_stats_manuscript"
date: "2025-07-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(jsonlite)
library(tidyverse)
library(purrr)
library(ggrepel)
library(here)
library(irr)
library(ggvenn)
library(gridExtra)
library(psych)
library(lme4)
library(stringr)
library(lmerTest)  # For p-values in lme4 models
library(papaja) # For APA style tables and figures
source(here("helper.R"))
```

#library(skimr)     # For quick and nice summaries
#library(performance) # For model diagnostics

# Introduction
Read in language data. 

```{r}
word_language_mappings <- read_csv(here("data/ratings", "alllangs.csv")) %>% 
  filter(!is.na(uni_lemma))  %>% 
  mutate(word = uni_lemma) %>%
  select(word, language)
```
# Study 1A

```{r}
# english_sample1_A <- read_csv(here("data/ratings","english_sample1.csv")) # all dimensions
# english_sample1_B <- read_csv(here("data/ratings", "english_sample1_redo.csv")) # all dimensions
# english_sample1_C <- read_csv(here("data/ratings","english_sample1_redo2.csv")) # all dimensions
# 
# english_sample1 <- english_sample1_A %>% 
#   full_join(english_sample1_B) %>%
#   full_join(english_sample1_C) %>% distinct()


english_sample1 <- read_csv(here("data/ratings","correct_english_sample1.csv")) # only category

english_sample1  <- english_sample1 %>% filter(trial_type == "survey-multi-choice")%>%
  mutate(uni_lemma = theword, word = theword) %>%
  filter(!str_detect(block, "practice") & !str_detect(block, "attention") ) %>% 
  filter(!is.na(response))


english_sample1 %>% group_by(theword, block) %>% summarize(count = n()) %>% filter(count <100) %>% ggplot(aes(x= count)) + geom_histogram()

#english_sample1$response <- sapply(english_sample1$response, extractValue)
#APPLY THE extractValue mapping function to the response column only when the response is following the expected JSON-like structure like this one "{\"Q0\":\"color\"}" 
english_sample1$response <- sapply(english_sample1$response, clean_json_response)


english_sample1$response_numeric <- mapply(map_categories, english_sample1$response, english_sample1$block)

english_sample1 <- english_sample1 %>% group_by(word, block) %>%
  mutate(totcount = n()) %>% ungroup()

english_sample1 %>% filter (word == "accident" & block == "category_organization")
english_sample1 %>% filter (word == "accident" & block == "solidity")


write_csv(english_sample1, here("data/predictors", "english_sample1_raw.csv"))

```

```{r}
# how many ratings per response per block

english_sample1_props <- english_sample1 %>%
  group_by(word, theword, block, response, totcount, uni_lemma, response_numeric) %>%
  summarise(count = n(), .groups = 'drop')

english_sample1_props <- english_sample1_props %>%
  mutate(proportion = count/totcount)

english_sample1_props <- merge(english_sample1_props, word_language_mappings, by = "word") %>% 
  filter( !is.na(word), !is.na(language))

english_sample1_props <- english_sample1_props %>% mutate(eng_response= response)

#write_csv(english_sample1_props, here("data/predictors", "english_sample1_props.csv"))

```

```{r}


study1a_data <- english_sample1 %>%
  filter(block == "category_organization") %>%
  group_by(word, theword, block, response, totcount, uni_lemma, response_numeric) %>%
  mutate(count = n(), .groups = 'drop')

study1a_data <- study1a_data %>%
  mutate(shape_rating = count/totcount) %>%
  filter(response == "shape")

study1a_data <- merge(study1a_data, word_language_mappings, by = "word") %>% 
  filter( !is.na(word), !is.na(language))

study1a_data <- study1a_data %>% mutate(eng_response= response)

write_csv(study1a_data, here("data/predictors", "study1a_data.csv"))

```

# Study 1B

```{r}

# english_sample2_B <- read_csv(here("data/ratings", "english_sample2_redo1.csv")) 
# english_sample2_C <- read_csv(here("data/ratings","english_sample2_redo3.csv")) 
# english_sample2_D <- read_csv(here("data/ratings","english_sample2_redo4.csv")) 
# 
# 
# english_sample2 <- english_sample2_A %>% 
#   full_join(english_sample2_B) %>%
#   full_join(english_sample2_C)  %>%
#   full_join(english_sample2_D) %>% 
#   distinct()

english_sample2 <- read_csv(here("data/ratings","correct_english_sample2.csv")) # only category

english_sample2 <- english_sample2 %>% 
  filter(!is.na(response)) %>% 
  filter(!is.na(word)) 

english_sample2 %>% group_by(theword, block) %>% summarize(count = n()) %>% filter(count <100) %>% ggplot(aes(x= count)) + geom_histogram()


english_sample2  <- english_sample2 %>% filter(trial_type == "survey-multi-choice")%>%
  mutate(uni_lemma = theword, word = theword) %>%
  filter(!str_detect(block, "practice") & !str_detect(block, "attention") ) %>% 
  filter(!is.na(word))

english_sample2$response <- sapply(english_sample2$response, clean_json_response)

english_sample2$response_numeric <- mapply(map_categories, english_sample2$response, english_sample2$block)

english_sample2 <- english_sample2 %>% group_by(word, block) %>%
  mutate(totcount = n()) %>% ungroup()

english_sample2 %>% filter(is.na(response))
```

```{r}
# how many ratings per response per block
english_sample2_props <- english_sample2 %>%
  group_by(word, theword, block, response, response_numeric ,totcount, uni_lemma) %>%
  summarize(count = n(), .groups = 'drop')

english_sample2_props %>% filter(is.na(response))

english_sample2_props <- english_sample2_props %>%
  mutate(proportion = count/totcount)

english_sample2_props %>% filter(is.na(response))


english_sample2_props <- merge(english_sample2_props, word_language_mappings, by = "word") %>%
  filter( !is.na(word), !is.na(language))

english_sample2_props %>% filter(is.na(response))


english_sample2_props <- english_sample2_props %>% mutate(eng_response= response)
#write_csv(english_sample2_props, here("data/predictors", "english_sample2_props.csv"))


```
```{r }
study1b_data <- english_sample2 %>%
  filter(block == "category_organization") %>%
  group_by(word, theword, block, response, totcount, uni_lemma, response_numeric) %>%
  mutate(count = n(), .groups = 'drop')
study1b_data <- study1b_data %>%
  mutate(shape_rating = count/totcount) %>%
  filter(response == "shape")
study1b_data <- merge(study1b_data, word_language_mappings, by = "word") %>%
  filter( !is.na(word), !is.na(language))
study1b_data <- study1b_data %>% mutate(eng_response= response)
write_csv(study1b_data, here("data/predictors", "study1b_data.csv"))

```


# Study 2A
```{r}
# korean data 
korean_sample1_A <- read_csv(here::here("data/first-sample-korean","korean_online.csv"))
korean_sample1_B <- read_csv(here::here("data/first-sample-korean","korean_replication_prolific.csv"))

korean_sample1_C <- read_csv(here::here("data/first-sample-korean","korean_unilemma_firstrun.csv"))

korean_sample1_D <- read_csv(here::here("data/first-sample-korean","koreanvarlogs2.csv"))

korean_sample1_E <- read_csv(here::here("data/first-sample-korean","korean_s1_sep25.csv"))

korean_sample1_F <- read_csv(here::here("data/first-sample-korean","korean_s1_sep28.csv"))

korean_sample1 <- korean_sample1_A %>% 
  full_join(korean_sample1_B) %>%
  full_join(korean_sample1_C) %>%
  full_join(korean_sample1_D) %>%
  full_join(korean_sample1_E) %>%
  full_join(korean_sample1_F) %>%
  distinct()

korean_sample1 %>% group_by(theword, block) %>% summarize(count = n()) %>% filter(count <100) %>% ggplot(aes(x= count)) + geom_histogram()


korean_sample1 <- korean_sample1 %>% 
  filter(trial_type == "survey-multi-choice")%>%
  mutate(word = theword) %>%
  filter(!str_detect(block, "practice") & !str_detect(block, "attention") ) %>% 
  filter(!is.na(word)) %>% 
  filter(!is.na(response))

korean_sample1$response <- sapply(korean_sample1$response, extractValue)

korean_sample1 %>% filter(is.na(response))


```


translate responses to english
```{r}
# translate responses to english using these responses: > unique(korean_data$response) [1] "고형물"                    "셀 수 있다"          [3] "셀 수 없다"                "물질"                     [5] "해당 사항 없음"            "형태"                      [7] "비고형물"                  "색상"                     [9] "잘 모르겠다/구분이 어렵다"

korean_sample1$eng_response <- sapply(korean_sample1$response, function(x) {
  if (x == "고형물") {
    return("solid")
  } else if (x == "셀 수 있다") {
    return("countable")
  } else if (x == "셀 수 없다") {
    return("uncountable")
  } else if (x == "물질") {
    return("substance")
  } else if (x == "해당 사항 없음") {
    return("not applicable")
  } else if (x == "형태") {
    return("form")
  } else if (x == "비고형물") {
    return("non-solid")
  } else if (x == "색상") {
    return("color")
  } else if (x == "잘 모르겠다/구분이 어렵다") {
    return("not sure")
  } else {
    return(x)
  }
})
    
  
```

```{r}
korean_unilemma = read_csv(here("data/ratings", "koreanwords.csv")) %>%
  mutate(theword = item_definition)

korean_sample1 <- korean_sample1 %>%
  merge(korean_unilemma %>% select(theword, uni_lemma), by = "theword") %>% filter(!is.na(uni_lemma)) %>% distinct()

```

```{r }
korean_sample1 <- korean_sample1 %>% group_by(word, block, uni_lemma, theword) %>%
  mutate(totcount = n()) %>% ungroup()

# how many ratings per response per block
korean_sample1_props <- korean_sample1 %>%
  group_by(word, block, response, totcount, uni_lemma, eng_response, theword) %>%
  summarize(count = n(), .groups = 'drop')

korean_sample1_props <- korean_sample1_props %>%
  mutate(proportion = count/totcount, language= "Korean", matching_response = case_when(eng_response == "form" ~ "shape",
                            TRUE ~ eng_response))

write_csv(korean_sample1_props, here("data/predictors", "korean_sample1_props.csv"))
  
```

# Study 2B
```{r }
korean_sample2_props <- read_csv(here("data/ratings", "korean_sample2.csv")) # chosun

korean_sample2_props <- korean_sample2_props %>% 
  mutate(uni_lemma = Word, language = "Korean", eng_response = response,
         matching_response = case_when(response == "Form" ~ "shape",
                                       TRUE ~ response)) %>% 
  filter(!is.na(response))


```

unique(korean_sample1_props$eng_response)
[1] "form"           "countable"      "solid"         
[4] "substance"      "non-solid"      "not applicable"
[7] "color"          "uncountable"    "not sure" 

unique(korean_sample2_props$eng_response)
 [1] "Form"                                 
 [2] "Form, Color"                          
 [3] "Form, Color, Material"                
 [4] "Form, Material"                       
 [5] "Material"                             
 [6] "Not applicable"                       
 [7] "Can be counted"                       
 [8] "Cannot be counted"                    
 [9] "I don't know/It's hard to say"        
[10] "Not a solid"                          
[11] "Solid"                                
[12] "Color"                                
[13] "Color, Material"                      
[14] "Form, Color, Material, Not applicable"
[15] "Form, Color, Not applicable"          
[16] "Material, Not applicable"             
[17] "Form, Material, Not applicable"       
[18] "Color, Not applicable"     


unique(english_sample1_props$eng_response)
 [1] "material"        "none of these"  
 [3] "count noun"      "mass noun"      
 [5] "unclear/unknown" "non-solid"      
 [7] "shape"           "solid"          
 [9] "color"           NA   
 
 unique(english_sample2_props$eng_response)
[1] "none of these" "material"      "shape"        
[4] "color"         NA   

unifying responses:
``` {r}

# Function to unify responses across datasets

unify_responses <- function(df) {
  df %>%
    mutate(
      cleaned_response = str_to_lower(eng_response),
      cleaned_response = str_replace_all(cleaned_response, "[^a-z\\s]", ""),
      cleaned_response = str_squish(cleaned_response),

      matching_response = case_when(
        # MULTIPLE responses
        str_detect(eng_response, ", ") ~ "multiple",

        # SHAPE-related
        str_detect(cleaned_response, "form|shape") ~ "shape",

        # MATERIAL-related
        str_detect(cleaned_response, "material|substance") ~ "material",

        # COLOR-related
        str_detect(cleaned_response, "color") ~ "color",

        # SOLIDITY
        str_detect(cleaned_response, "(?<!not a )\\bsolid\\b") ~ "solid",
        str_detect(cleaned_response, "non[- ]?solid|not a solid|non-solid|nonsolid") ~ "non-solid",

        # COUNTABILITY
        str_detect(cleaned_response, "\\bcountable\\b|\\bcount noun\\b|\\bcan be counted\\b") ~ "count",
        str_detect(cleaned_response, "\\buncountable\\b|\\bmass noun\\b|\\bcannot be counted\\b") ~ "mass",

        # UNKNOWN/OTHER
        str_detect(cleaned_response, "none of these|not applicable") ~ "none of these",
        str_detect(cleaned_response, "not sure|dont know|unknown|unclear|hard to say") ~ "unknown",

        # If nothing matches
        TRUE ~ "other"
      )
    )
}


```

```{r}
# Apply the unifying function to each dataset
korean_sample1_props <- unify_responses(korean_sample1_props)
korean_sample2_props <- unify_responses(korean_sample2_props)
english_sample1_props <- unify_responses(english_sample1_props)
english_sample2_props <- unify_responses(english_sample2_props)

```

```{r}
korean_sample1_props$eng_response %>% unique()

korean_sample1_props$cleaned_response %>% unique()

korean_sample1_props$matching_response %>% unique()
```

```{r}
korean_sample2_props$eng_response %>% unique()

korean_sample2_props$cleaned_response %>% unique()

korean_sample2_props$matching_response %>% unique()

korean_sample2_props %>% filter(eng_response == "Not a solid") %>% select(eng_response, cleaned_response, matching_response) %>% unique()

```

```{r}
english_sample1_props$matching_response %>% unique()
english_sample2_props$matching_response %>% unique()

```

```{r}
write_csv(english_sample2_props, here("data/predictors", "english_sample2_props.csv"))
write_csv(english_sample1_props, here("data/predictors", "english_sample1_props.csv"))
write_csv(korean_sample1_props, here("data/predictors", "korean_sample1_props.csv"))
write_csv(korean_sample2_props, here("data/predictors", "korean_sample2_props.csv"))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
