
# English Study 1

```{r message=FALSE, warning=FALSE}
english_sample1 <- read_csv(here("data/predictors","english_sample1_props.csv")) 
english_sample1 <- english_sample1 %>% filter(!is.na(response)) %>% distinct()
english_sample1_raw <- read_csv(here("data/predictors","english_sample1_raw.csv")) 
# Apply the mapping function in helper.R to the response column
english_sample1$response_numeric <- mapply(map_categories, english_sample1$response, english_sample1$block)
# more 2 participants -- 21 words less than 8 ratings

correspondence <- english_sample1 |> filter (matching_response %in% c("solid", "count" ,"shape"))

```

```{r , echo=TRUE}
# how much correspondence there is between the three dimensions 
correspodence_data <-  english_sample1 %>% select(-c(language)) %>%
  distinct() %>%
  filter(response %in% c("solid", "shape", "count noun")) %>%  # Filter only the desired responses
  select(word, response, proportion) %>%  # Keep only relevant columns
  pivot_wider(names_from = response, values_from = proportion, values_fill = 0)  # Reshape

correspodence_data %>% select(`count noun`, shape, solid) %>% corr.test()

```

```{r agreement_plot, include=TRUE}
consensus_data <- english_sample1_raw %>%
  group_by(block, uni_lemma, word) %>%
  summarise(
    mode_rating = get_mode(response_numeric),
    agreement_score = (mean(response_numeric == mode_rating, na.rm = TRUE) -0.009) * 100,
    sd_rating = sd(response_numeric, na.rm = TRUE)
  )

ggplot(consensus_data, aes(x = sd_rating, y = agreement_score, color = block)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Agreement Score vs. Rating Standard Deviation", x = "Standard Deviation of Ratings", y = "Agreement Score (%)") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3")
```

# English Study 2
```{r }
english_sample2 <- read_csv(here("data/predictors","english_sample2_props.csv"))
english_sample2 <- english_sample2 %>% filter(!is.na(response)) %>% distinct()
english_sample2_raw <- read_csv(here("data/predictors","english_sample2_raw.csv"))
# Apply the mapping function in helper.R to the response column
english_sample2$response_numeric <- mapply(map_categories, english_sample2$response, english_sample2$block)

# more 1 or 2 participants -- 29 words less than 8 ratings

# english_sample1 %>% filter(totcount < 8, block == "category_organization") %>% select(word) %>% distinct() %>% jsonlite::toJSON()

#english_sample2 %>% filter(totcount < 8, block == "category_organization") %>% select(word) %>% distinct() %>% jsonlite::toJSON()

```

# Reliability analysis 

```{r }
eng_samp1_shape <- english_sample1 %>% filter(response == "shape") %>% select(word, proportion)
eng_samp2_shape <- english_sample2 %>% filter(response == "shape") %>% select(word, proportion)

# Merge the data frames based on 'word_id'
merged_data <- merge(eng_samp1_shape, eng_samp2_shape, by = "word", suffixes = c(".sample1", ".sample2")) %>% distinct()
```

## split-half estimate of reliability
through simple pearson correlation 
```{r}
# Calculate the correlation between proportion values
correlation_score <- cor(merged_data$proportion.sample1, merged_data$proportion.sample2, use = "complete.obs") # Handles missing values

# Print the correlation score
print(paste("correlation between shape proportions in 1A VS 1B is: ", correlation_score))

```

```{r }
ggplot(merged_data, aes(x = proportion.sample1, y = proportion.sample2)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Item-Level Agreement: Shape Ratings in Study 1A vs 1B")
```

## overall distribution
```{r }
# Load your data (adjust file paths as needed)
study1a_data <- read_csv(here("data/predictors","study1a_data.csv"))

#fix names of columns
study1a_data <- study1a_data %>%
  rename(participant_id = `_id`)

study1b_data <-  read_csv(here("data/predictors","study1b_data.csv"))
study1b_data <- study1b_data %>%
  rename(participant_id = `_id`)


# 1. DATA PREPARATION
# ===================
study1a_data<- study1a_data %>% 
    select(uni_lemma, language, shape_rating) %>%
    mutate(study = "1A")

study1b_data <- study1b_data %>% 
    select(uni_lemma, language, shape_rating) %>%
    mutate(study = "1B")

# Combine datasets with study identifier
combined_data <- bind_rows( 
  study1b_data, study1a_data
   ) %>%
  filter(!is.na(shape_rating)) %>%  # Remove missing values
mutate(shape_rating = as.numeric(shape_rating)) %>%
  distinct()  # Remove duplicates

combined_data %>% filter(study == "1A")
combined_data %>% filter(study == "1B")

# Create wide format for reliability analysis
ratings_wide <- combined_data %>%
  pivot_wider(names_from = study, 
              values_from = shape_rating,
              names_prefix = "Study_") %>%
  filter(!is.na(Study_1A) & !is.na(Study_1B))  # Only words rated in both studies

```

Use a Kolmogorov–Smirnov (KS) test to see if the two distributions differ significantly.
```{r }
ks.test(combined_data$shape_rating[combined_data$study == "1A"],
        combined_data$shape_rating[combined_data$study == "1B"])

```

A classic example of low item-level agreement but high aggregate-level agreement.

It doesn’t necessarily mean the measure is unreliable — instead, it may suggest:

That there's measurement noise at the item level (e.g., variability in how different raters interpret the words).

Or that your construct is more meaningful at the group level — i.e., shape ratings are better thought of as a property of categories or populations, not precise values for individual words.

In psychometrics, this distinction often arises between individual-level reliability (e.g., inter-rater reliability for specific items) and scale-level consistency (e.g., internal consistency across a battery of items).

## ICC

cor()) tells how well rank order is preserved. Intraclass Correlation Coefficient (ICC) tells how much of the variance in scores is due to between-item differences vs. noise. Useful for assessing rater consistency or measurement reliability.
```{r }
library(irr)
icc_result <- icc(data.frame(merged_data$proportion.sample1, merged_data$proportion.sample2), model = "twoway", type = "consistency", unit = "single")
print(icc_result)

```

## Bland-Altman Plot
visualize agreement between two measurements (especially useful when comparing two raters or samples). shows the mean difference between ratings and any bias or systematic shift across rating ranges

```{r }
if (!requireNamespace("BlandAltmanLeh", quietly = TRUE)) {
      install.packages("BlandAltmanLeh")
}
library(BlandAltmanLeh)

bland.altman.plot(merged_data$proportion.sample1, merged_data$proportion.sample2)

```

## Prophecy: spearman-brown prediction formula 
``` {r}
predicted_reliability <- (2 * correlation_score) / (1 + correlation_score)
```
we examined the measurement properties by measuring reliability, then examine the 


## Using pearson correlation between two halves of the data (split-half reliability)

## Internal consistency (Cronbach's alpha) for each dimension in each study

## Item-Level Error Inspection

```{r}
merged_data$abs_diff <- abs(merged_data$proportion.sample1 - merged_data$proportion.sample2)
summary(merged_data$abs_diff)
# Identify items with the largest discrepancies
top_discrepancies <- merged_data %>% arrange(desc(abs_diff)) %>%
  head(15)
print(top_discrepancies)
```

# Korean 

```{r }
korean_sample1 <- read_csv(here("data/predictors","korean_sample1_props.csv"))
korean_sample2 <- read_csv(here("data/predictors","korean_sample2_props.csv"))
# filter the words fish and window, which are homonyms in Korean and have different meanings in English, to avoid confusion in the analysis
korean_sample1 <- korean_sample1 %>% filter(uni_lemma != "fish", uni_lemma != "window")
korean_sample2 <- korean_sample2 %>% filter(uni_lemma != "fish", uni_lemma != "window", Word != "Fish", Word != "Window", Word != "fish", Word != "window")


english_sample1_koreanlist <- english_sample1 %>% filter (language == "Korean")

words <- read_csv(here("data/ratings", "alllangs.csv")) %>% 
  filter(!is.na(uni_lemma))  %>% 
  mutate(word = uni_lemma) %>%
  select(word, language, category)

korean_words <- read_csv(here("data/ratings", "alllangs.csv")) %>% 
  filter(!is.na(uni_lemma), language == "Korean")  %>% 
  mutate(word = item_definition, theword = item_definition) %>% 
  select (word, theword, uni_lemma, language, category)

korean_words <- korean_words %>% 
  mutate(uni_lemma = str_to_lower(uni_lemma))

korean_sample2 <- korean_sample2 %>% 
  mutate(Word = str_to_lower(Word))

korean_sample2 <- korean_sample2 |> merge(korean_words, 
        by = "theword") |> 
  select(-ends_with(".x")) %>% distinct() %>%
  rename_with(~ str_remove_all(., "\\.y$"), everything())

```

```{r }
canonical_responses <- c(
  "shape", "material", "color",
  "count", "mass",
  "solid", "non-solid",
  "none", "unknown",
  "multiple"
)

response_map <- tribble(
  ~block,                  ~raw_response,        ~response_norm,

  # -------------------------
  # category_organization
  # -------------------------
  "category_organization", "shape",              "shape",
  "category_organization", "color",              "color",
  "category_organization", "substance",           "material",
  "category_organization", "material",            "material",
  "category_organization", "multiple",            "multiple",
  "category_organization", "none of these",       "none",
  "category_organization", "not applicable",      "none",

  # -------------------------
  # solidity
  # -------------------------
  "solidity",              "solid",              "solid",
  "solidity",              "non-solid",          "non-solid",
  "solidity",              "unknown",            "unknown",
  "solidity",              "unclear/unknown",    "unknown",
  "solidity",              "not applicable",     "unknown",

  # -------------------------
  # count_mass
  # -------------------------
  "count_mass",            "countable",          "count",
  "count_mass",            "count",              "count",
  "count_mass",            "count noun",         "count",
  "count_mass",            "uncountable",        "mass",
  "count_mass",            "mass",               "mass",
  "count_mass",            "mass noun",          "mass",
  "count_mass",            "not sure",           "unknown",
  "count_mass",            "unknown",            "unknown"
)

```

```{r }
korean1_norm <- korean_sample1 %>%
  left_join(
    response_map,
    by = c("block", "matching_response" = "raw_response")
  )

korean2_norm <- korean_sample2 %>%
  left_join(
    response_map,
    by = c("block", "matching_response" = "raw_response")
  )

unique(c(korean1_norm$response_norm,
         korean2_norm$response_norm))

english1_norm <- english_sample1_koreanlist %>%
  left_join(
    response_map,
    by = c("block", "response" = "raw_response")
  )


english2_norm <- english_sample2 %>% filter(language == "Korean") %>%
  left_join(
    response_map,
    by = c("block", "matching_response" = "raw_response") 
  )

setdiff(
  unique(c(english1_norm$response_norm,
           english2_norm$response_norm)),
  canonical_responses
)

```
```{r }

kor1 <- korean1_norm %>% 
    select(uni_lemma, response_norm, block, proportion) %>% 
    filter(response_norm %in% c("shape", "count", "solid")) %>%
    mutate(sample = "kor1") %>% distinct()

kor1 %>% filter(uni_lemma == "bag")

kor2 <- korean2_norm %>% 
    select(Word, response_norm, block, proportion) %>%
    filter(response_norm %in% c("shape", "count", "solid")) %>%
    mutate(sample = "kor2") %>% distinct() %>%
  rename(uni_lemma = Word)

kor2 %>% filter(uni_lemma == "bag")

korean_all <- bind_rows(kor1, kor2) %>% distinct()

korean_all %>% filter(uni_lemma == "bag", response_norm == "shape", block == "category_organization")

korean_all %>%
  summarise(n = n(), .by = c(uni_lemma, response_norm, block, sample)) %>%
  filter(n > 1)

english_all <- bind_rows(
  english1_norm %>% select(uni_lemma, response_norm, block, proportion) %>%
    filter(response_norm %in% c("shape", "count", "solid")) %>%
    mutate(sample = "eng1"),
  english2_norm %>%  select(uni_lemma, response_norm, block, proportion) %>%
    filter(response_norm %in% c("shape", "count", "solid")) %>%
    mutate(sample = "eng2")
) %>% distinct()

sample_map <- tribble(
  ~korean_sample, ~english_sample,
  "kor1",         "eng1",
  "kor2",         "eng2"
)

```

```{r }

all_data <- bind_rows(english_all, korean_all) 

wide_data <- all_data %>%
  pivot_wider(
    names_from = sample,
    values_from = proportion
  ) 

all_data %>%
  summarise(n = n(), .by = c(uni_lemma, response_norm, block, sample)) %>%
  filter(n > 1)


icc_data <- wide_data %>%
  select(eng1, eng2, kor1, kor2)
icc_data
psych::ICC(icc_data)

write_csv(wide_data, here("data/predictors", "reliability_wide_alldata.csv"))
write_csv(all_data, here("data/predictors", "reliability_long_alldata.csv"))


```
Reliability of item-level proportions across norming samples was assessed using intraclass correlation coefficients (ICCs; McGraw & Wong, 1996). Treating samples as fixed effects, average-measures reliability was high, ICC(3,4) = .87, 95% CI [.86, .89], indicating strong consistency across English and Korean norming datasets.

If you want the random-effects framing: Average-measures reliability remained high under a random-effects model, ICC(2,4) = .83, 95% CI [.72, .88].

A single norming sample gives you a moderately reliable estimate of the underlying proportion. If you average across samples, reliability is high to excellent, suggesting that the item-level proportions are stable and consistent across different norming samples and languages > ICC2


```{r }
wide_data %>%
  select(eng1, eng2, kor1, kor2) %>%
  corr.test()

```
# Language-aware ICC
```{r }
# “After obtaining the most reliable estimate within each language, do items show similar structure across languages?”

long_data <- all_data %>%
  mutate(
    language = case_when(
      sample %in% c("eng1", "eng2") ~ "English",
      sample %in% c("kor1", "kor2") ~ "Korean"
    )
  )

lang_means <- long_data %>%
  group_by(uni_lemma, response_norm, block, language) %>%
  summarise(proportion = mean(proportion), .groups = "drop")

lang_means_wide <- lang_means %>%
  pivot_wider(names_from = language, values_from = proportion)

cor(lang_means_wide$English, lang_means_wide$Korean, use = "pairwise")
```
When we get the average within language, items are ranked very similarly in English and Korean.
Reliability was assessed both within and across languages. Within each language, item-level shape proportions were moderately to strongly correlated across the two independent norming samples (Korean: r = .38; English: r = .54). Applying the Spearman–Brown correction indicated that averaging across samples substantially improved reliability (Korean: .55; English: .70). Accordingly, language-specific averages were used for subsequent analyses. To assess cross-language consistency, we correlated item-level proportions averaged within English and Korean. This analysis revealed a strong cross-language association (r = .71), indicating that items exhibit highly similar structure across languages.

If a reviewer asks “why not ICC?”, your answer is simple:

We used split-half correlations with Spearman–Brown correction because they directly quantify the reliability of averaged item-level proportions and make minimal assumptions about variance homogeneity. see below for a more detailed explanation of the variance components and ICC approach.

```{r }
# After averaging within language, how much variance is due to items vs languages vs noise?
# a variance-partitioning model, not a reliability coefficient

m <- lmer(
  proportion ~ 1 +
    (1 | uni_lemma) +
    (1 | language),
  data = lang_means,
  REML = TRUE
)

variance_components <- as.data.frame(VarCorr(m))
variance_components$proportion <- variance_components$vcov / sum(variance_components$vcov)
vc <- variance_components
vc
# 1. “How reliable are samples within the same language?”
ICC_within_lang <-
  vc$vcov[vc$grp == "uni_lemma"] /
  (vc$vcov[vc$grp == "uni_lemma"] +
   vc$vcov[vc$grp == "language"] +
   vc$vcov[vc$grp == "Residual"])
ICC_within_lang

# 2. “How reliable are samples across languages?” “How much variance is item-level vs language-driven?”
ICC_cross_lang <-
  vc$vcov[vc$grp == "uni_lemma"] /
  sum(vc$vcov)
ICC_cross_lang

```
| Component   | Variance | Proportion |
| ----------- | -------- | ---------- |
| `uni_lemma` | 0.0144   | **19.4%**  |
| `language`  | 0.0043   | **5.8%**   |
| Residual    | 0.0557   | **74.8%**  |

About 19% of total variance in language-averaged proportions is attributable to stable item differences.
not low, given that:
* the DV is a bounded proportion
* many items cluster near 0 or 1
* residual variance includes binomial + task noise

The short, accurate meaning of “residual variance includes binomial + task noise” = 
After accounting for stable item differences and language effects, the remaining unexplained variability reflects (a) sampling variability inherent to proportions and (b) noise introduced by the experimental task itself.

Those two sources are structural, not mistakes.
1️⃣ Binomial variance (the unavoidable part)

the dependent variable is a proportion:
Even if an item has a true underlying probability p, the observed proportio nmust vary across samples because participant responses are discrete.

For a binomial process:
Variance is largest near p = .5
Variance shrinks near 0 or 1
Smaller n → more variance

This variance exists even if participants are perfectly consistent.
mixed model treats all of this as residual variance.

2️⃣ Task noise (human + procedural variability)

Beyond binomial sampling, participants also differ in ways that are not item-stable:
Examples:
momentary attention
interpretation of instructions
response biases
fatigue
interface effects
trial context
These add variability that:
is not systematic across items
does not replicate across samples
does not reflect “true” item properties
This is task noise.

3️⃣ Why both end up in the residual
model: proportion ~ 1 + (1 | uni_lemma) + (1 | language)

has no term that can explain:
trial-level randomness
participant-level variability
binomial sampling error
So all of that variance gets absorbed into:
Residual
This is expected.

4️⃣ Why this matters for interpretation

This is the crucial insight:
A large residual variance does NOT mean your data are unreliable.
It means:
your measurement is noisy at the participant level
but noise cancels out when you average
which is exactly why reliability improves with aggregation
That’s why you saw:
moderate single-sample correlations
high reliability after averaging

strong cross-language correlations
All three are compatible.

5️⃣ Why proportions make residuals look especially big

Two reasons:

A. Bounded scale
Proportions are restricted to 
[0,1] which:
compresses between-item variance
inflates the relative size of residual variance

B. Gaussian mismatch

we're fitting a Gaussian mixed model to binomially generated data.
This doesn’t invalidate the model, but it:
shifts binomial variance into the residual
lowers variance-based “ICC-like” ratios
That’s why variance-partition ICCs often look “low” for proportions.

 to model this explicitly: could separate binomial variance from task noise by using a binomial or beta mixed model:

glmmTMB(
  cbind(k, n - k) ~ 1 + (1 | uni_lemma) + (1 | language),
  family = binomial,
  data = ...
)


The residual variance reflects unavoidable binomial sampling variability in proportion estimates as well as unsystematic task-related noise, rather than unreliability of the underlying item-level structure.

At the word level, we asked whether the words that strongly elicit count responses in the countability block are the same words that also strongly elicit solid and shape response relative to their respective alternatives. Pearson correlation analysis revealed a moderate negative correlation between countability and both solidity and shape rating. This could also be explained by the carryover effect between blocks, where participants could use the global context of the previous blocks to inform their ratings in the following blocks creating a contrast effect. For example, if a word is strongly rated as countable, participants may be less likely to rate it as solid or shape-based. On the other hand, sample 2A showed a positive correlation because the randomization of blocks and words could have led to more independent ratings across blocks. 

# old code
```{r }
response_map <- tribble(
  ~block,                    ~korean,           ~english_ref,
  # category_organization
  "category_organization",   "substance",        "material",
  "category_organization",   "not applicable",   "none of these",
  "category_organization",   "shape",            "shape",
  "category_organization",   "color",            "color",

  # solidity
  "solidity",                "solid",            "solid",
  "solidity",                "non-solid",        "non-solid",
  "solidity",                "not applicable",   "unknown",

  # count_mass
  "count_mass",              "countable",        "count",
  "count_mass",              "uncountable",      "mass",
  "count_mass",              "not sure",         "unknown"
)

korean_norm <- korean_sample1 %>%
  left_join(
    response_map,
    by = c("block" = "block", "matching_response" = "korean")
  ) %>%
  mutate(response_norm = english_ref)

english_norm <- english_sample1_koreanlist %>%
  mutate(
    response_norm = case_when(
      response %in% c("unclear/unknown", "unknown") ~ "unknown",
      TRUE ~ response
    )
  )

english_norm <- english_norm %>%
  mutate(
    response_norm = case_when(
      response %in% c("count", "count noun") ~ "count",
      response %in% c("mass", "mass noun") ~ "mass",
      response %in% c("unclear/unknown", "unknown") ~ "unknown",
      response == "none of these" ~ "none",
      TRUE ~ response
    )
  )

korean_norm <- korean_norm %>%
  mutate(
    response_norm = case_when(
      response_norm == "none of these" ~ "none",
      TRUE ~ response_norm
    )
  )

setdiff(unique(korean_norm$response_norm),
        unique(english_norm$response_norm))

```

```{r }
merged <- korean_norm %>%
  select(uni_lemma, block, response_norm, proportion) %>%
  left_join(
    english_norm %>%
      select(uni_lemma, block, response_norm,
             proportion_eng = proportion),
    by = c("uni_lemma", "block", "response_norm")
  ) %>%
  distinct()

merged %>%
  filter(uni_lemma == "bag", block == "count_mass")

merged_wide <- merged %>%
  mutate(
    dimension = case_when(
      block == "count_mass" & response_norm == "count" ~ "count",
      block == "category_organization" & response_norm == "shape" ~ "shape",
      block == "solidity" & response_norm == "solid" ~ "solid",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(dimension)) %>%
  select(uni_lemma, dimension, proportion, proportion_eng) %>%
  pivot_wider(
    names_from = dimension,
    values_from = c(proportion, proportion_eng)
  )

merged_wide %>%
  select(
    proportion_count, proportion_eng_count,
    proportion_shape, proportion_eng_shape,
    proportion_solid, proportion_eng_solid
  ) %>%
  corr.test()

list(
  count = corr.test(merged_wide$proportion_count,
                    merged_wide$proportion_eng_count),
  shape = corr.test(merged_wide$proportion_shape,
                    merged_wide$proportion_eng_shape),
  solid = corr.test(merged_wide$proportion_solid,
                    merged_wide$proportion_eng_solid)
)


```